# Wrangle data for ASC RUM

## Set up
```{r setup}
    # knitr options
    knitr::opts_chunk$set(warning = FALSE, message = TRUE, echo = FALSE, fig.align =
                            'center', fig.width = 10, fig.height = 7) 
    
    # libraries 
    library(tidyverse)
    library(dplyr)
    library(ggplot2)
    library(sp)
    library(raster)
    library(rgeos)
    library(rgdal)
    library(sf)
    library(lwgeom)
    library(googledrive)
    library(units)
    library(nngeo)
    library(magrittr)
    library(todor)
    
    # data
    # BR <- st_read("./data/shp/natBR.shp") # boat ramps
    BRtrips <- readRDS("./data/gpkg/2.1_BRtrips.gpkg") %>% # from BRtrips.R
      st_transform(crs = 4283)
    
    npz <- st_read("./data/gpkg/ntz.gpkg") %>% 
      st_transform(crs = 4283)
    
    grid_base <- st_read("./data/gpkg/ascBase_v2.gpkg") %>% 
      st_transform(crs = 4283)
    
    dat <- read_csv("data/01_data/1.2_dat.csv")
    
    # mp <- st_read("./data/gpkg/pmmp_4326.gpkg") %>% 
    #   st_transform(crs = 4283)
    
    coast <- st_read("./data/gpkg/coast.gpkg") %>% 
      st_transform(crs = 4283)
    
    # simulation
    sim_name <- "sim1"       #'[#NOTE: check simulation name]  
    
    sim <- st_read(paste0("./data/gpkg/sims/asc_", sim_name, ".gpkg")) %>% 
      st_transform(crs = 4283)
    
    # functions
    source("./functions/spatialFunc.R") # spatial functions
    source("./functions/ascFunc.R") # function to make asc grid
    source("./functions/theme.R") # plot theme
    
    sf_use_s2(FALSE)
```

## Data prep

All data filters have to be done before making the choice set.

```{r}
    dat %<>% 
  #'[#NOTE check all filters have been applied to data eg. activity type, site type and site]
  
      filter(ActivityType == "Extractive" & SiteType == "Boat" & str_detect(Site, 'BR')) %>% # filtering activity
      filter(!is.na(UseLat), !is.na(UseLong)) %>% # filtering empty coords
      st_as_sf(coords = c("UseLong", "UseLat"), crs = 4283)  ## make sf object
      
    # need one trip per row (longest site visited on trip)
    dat %<>% 
      group_by(PersonID, TripID) %>% 
      slice(which.max(decDuration)) %>% 
      ungroup() 
      
      dat <- st_crop(dat, grid_base) # removing outliers 
    
      #'[#VALIDATE: TRUE, no duplicated trip IDs)]
      identical((which(duplicated(dat$TripID) == TRUE)), integer(0)) == TRUE
```

## ASC grid and simulation stats

Creates smallest ASC grid possible within the base grid, given the distribution of data. Each cell must have at least 1 data point in it. 
Calculates basic stats for simulation and current marine park (eg. area, beach access etc.)

```{r}

deSliver <- function(x, threshold) {
  backToSf <- FALSE
  if (class(x)[1] == "sf") {
    backToSf <- TRUE
    x$tempArea <- as.numeric(st_area(x))
  } else {
    x<- st_as_sf(x)
    x$tempArea <- as.numeric(st_area(x))
  }

  #determine slivers by area
  xSlivers <- x[x$tempArea < threshold, ]
  xNotSlivers <- x[x$tempArea > threshold, ]
  if (nrow(xNotSlivers) < 1) {
    stop("Threshold exceeds the area of every polygon. Please select a smaller number")
  }

  #Split slivers from multipolygon, or nearest feature may be incorrect
  xSlivers <- suppressWarnings(st_cast(xSlivers, 'POLYGON'))

  #Find nearest non-sliver
  nearestFeature <- st_nearest_feature(xSlivers, xNotSlivers)

  #Merge each sliver polygon into nearest neighbour
  mergeSlivers <- lapply(
    unique(nearestFeature),
    FUN = function(i,
                   ns = xNotSlivers,
                   s = xSlivers,
                   nf = nearestFeature) {
      featurePolys <- nearestFeature == i
      xMerge <- s[featurePolys, ] %>%
        st_union(.)
      yMerge <- ns[i, ]
      #convert slivers back to multipolygon
      out <- sf::st_union(x = xMerge, y = yMerge) %>%
        as_Spatial(.)
      yMergeSpd <- sf::as_Spatial(yMerge)
      out <- SpatialPolygonsDataFrame(Sr = out,
                                 data = yMergeSpd@data,
                                 match.ID = FALSE)

      return(out)
    }
  )

  if (length(mergeSlivers) > 1) {
  m <- bind(mergeSlivers)
  } else {
    m <- mergeSlivers[[1]]
  }

  #Remove the temporary column
  m$tempArea <- NULL

  #remove self-intersecting geometries
  m <- gBuffer(spgeom = m, byid = TRUE, with = 0)

  if (backToSf) {
    m <- st_as_sf(m)
  }
  return(m)
}
```


```{r message = FALSE}
# asc grid
    grid <- ASCgrid(poly = grid_base, sz_current = npz, sz_sim = sim, point = dat, from = 0.03,
                    by = 0.02, crs = 4283, vert = FALSE) 

# manually merged data in QGIS
 # ascGrid for QGIS
    # st_write(grid,
    #          paste0("./data/gpkg/sims/2.1_ascGridQ_edit", sim_name, "_4283.gpkg"),
    #          append = FALSE) # need to write with st_write to open in QGISRUMs

# manual is only adding in id var - need rest of variables
# grid <- st_read("./data/gpkg/ascBase_v2_edit.gpkg") %>%
#       st_transform(crs = 4283)


# no new mp boundary to to run this
# sim_stats
    # sim_stats <- sim_stats(ascGrid = grid, point = dat, mp = mp, coast = coast, crs = 4283, sim_name = sim_name, path = "./plots/ASC RUM", width = 8, height = 11)
    # sim_stats
    
# mp_stats
    # mp_name <- "marmion"
    # mp_old <- st_read("./data/gpkg/mar_oldextent.gpkg") %>% st_transform(crs = 4283)
    # mp_stats <- mp_stats(ascGrid = grid, point = dat, mp = mp_old, coast = coast, crs = 4283, mp_name = "marmion", path = "./plots/ASC RUM", width = 8, height = 11)
    # write_csv(mp_stats, paste0("./plots/ASC RUM/", mp_name,"_stats.csv"))
  
#'[#REVIEW: this part of the function almost works. although merge is removing border from pt cloast sz? and is also removing mangrove bay cell which has valid data in it]  
  # st_rook = function(a, b = a) st_relate(a, b, pattern = "F***1****") # find neighbour (nb) function
  # 
  #  grid %<>% mutate(grid_nb = as.character(st_rook(.)))  # find neighbours (nb) for each grid cell
  #   merge_from <- grid %>% filter(data_present == FALSE) #cells with no data that need merged
  #   grid0_nb_list <- st_intersects(merge_from, grid) %>% unlist() #id list of all empty grid nbs
  #   merge_to <- grid %>% filter(gridID_alt %in% grid0_nb_list, data_present == TRUE) # df of nb to merge to
  #   merge_to_id <- as.character(unique(merge_to$gridID_alt))
  #     
  #   grid %<>% 
  #     mutate(merge_to = ifelse(str_detect(grid_nb, merge_to_id), merge_to_id, 0)) %>% 
  #     mutate(gridID_alt = ifelse(merge_to > 0, merge_to, gridID_alt)) %>% 
  #     group_by(gridID_alt) %>% 
  #     summarize() 
  #   
g <- grid %>% filter(grid$id %in% 10)

ggplot() +
  # geom_sf(data = grid, aes(fill = data_present))
#   # # geom_sf(data = grid0_nb_id, fill = "red")+
#   # # geom_sf(data = merge_from, fill = "blue")
  # geom_sf(data = coast_lam)+
  # geom_sf(data = poly_lam, fill = "red")
  # geom_sf(data = npz_lam)
  # geom_sf(data = a, fill = "red")
  # geom_sf(data = npz_lam)
  # geom_sf(data = g, aes(fill = as.factor(id)))
  # geom_sf(data = grid_base, fill = "yellow")+
  geom_sf(data = grid, aes(fill = as.factor(gridID_alt)))
  # geom_sf(data = npz, fill = "red")

# 24 is grid id of third false which is point cloast shore fishing zone
#18 is the mangrove bay zone that is mergeing for some reason
# b <- grid %>% filter( gridID_alt == 18)
  
# sim_lam <- sim %>%  st_transform(crs = 3112)
# npz_lam <- npz %>%  st_transform(crs = 3112)
# coast_lam <- coast %>%  st_transform(crs = 3112)
# poly_lam <- grid_base %>%  st_transform(crs = 3112)
# 
# b <- st_snap(sim_lam, npz_lam ,tolerance =  0.001)
# c <- st_snap(npz_lam, coast_lam ,tolerance =  0.001)
# 
# a <- st_difference(npz_lam, st_combine(poly_lam))
# 
# st_erase = function(x, y) st_difference(x, st_union(st_combine(y)))
# 
# a <- st_erase(poly_lam, npz_lam)
# 
# u = st_union(rbind(nbrs.buff, one))
# 
# grid1 <- grid %>% st_cast("MULTIPOLYGON") %>% st_cast("POLYGON")
# grid1$area <- as.numeric(set_units(st_area(grid1), km^2))
# grid1 %<>% filter(area > 10)
# 
# sim$area <- st_area(sim)
# npz$area <- st_area(npz)
# 
# which(grid1$sz %in% 1,)
```


## Attributes


```{r}
 grid <-  ascAtt(ascGrid = grid, grid_id = "gridID_alt", point = dat, br = BRtrips, br_id = "RampID",
              crs = 4283)
```
## Data wrangle
```{r}
    # isolating grid ID to append to dat
    gridID_vis <- grid %>%
      dplyr::select(gridID_alt, asc_geom) %>%
      distinct()

    # allocating sites to grids
    dat  %<>% st_join(gridID_vis, left = T, join = st_intersects)

    # selecting attributes, and tidying
    dat %<>%
      rename(gridID_vis = gridID_alt, dat_geom = geometry) %>%
      #'[#NOTE: select vars for model] 
      #'Mandatory vars: TripID, PersonID, gridID_vis, Site, BoatLength  
      dplyr::select(TripID, PersonID, gridID_vis, Site, BoatLength) %>% 
      distinct() %>%
      mutate(RampID = ifelse(Site %in% "ExmouthBR", 85, NA),
             RampID = ifelse(Site %in% "CoralBayBR", 72, RampID),
             RampID = ifelse(Site %in% "BundegiBR", 73, RampID),
             RampID = ifelse(Site %in% "TantabiddiBR", 87, RampID)) %>%
      # mutate(UseLong = st_coordinates(.)[,1]) %>%
      # mutate(UseLat = st_coordinates(.)[,2]) %>%
      # dplyr::select(-Site) %>%
      as.data.frame()

    # check there are no missing gridID_vis
    # a <- dat[which(is.na(dat$gridID_vis)), ]
    # a
    # 
    # ggplot() +
    #   geom_sf(data = grid$asc_geom) +
    #   geom_sf(data = a$dat_geom)
    
    #'[#VALIDATE: TRUE, there are no unassigned gridIDs]
    
    identical((which(is.na(dat$gridID_alt) == TRUE)), integer(0)) == TRUE
    
    # join
    t1 <- nrow(dat) # copying original data test join

    dat <- inner_join(dat, grid, by = "RampID") 
    
    #'[#VALIDATE: TRUE, number of sites * number of trips == nrow(dat), unless sites have been removed eg. beach launching sites]
    
    t1*nrow(gridID_vis) == nrow(dat)
```
## Distance to BR

Distance from the boat ramp to the centroid of data points in each grid that launched from that boat ramp - if not takes centroid of grid cell. 


```{r message=FALSE}
#'[#NOTE:could wrap below bit in separate function - use_centroid and other could be site_centroid] 

    dat <- dat %>% 
      st_as_sf() # turn to sf object
    
    # get centroid of use points
    tmp <- dat %>%
      group_by(gridID_vis, RampID) %>%
      summarise(use_centroid = st_union(dat_geom)) %>%
      st_centroid() %>%
      ungroup() %>%
      mutate(id = row_number())
    
    # validation plot: is the use centroid in the middle of each set of uses per boat ramp
    # ggplot() +
    #   geom_sf(data = dat, aes(geometry = asc_geom)) +
    #   geom_sf(data = dat, aes(geometry = dat_geom, color = as.factor(RampID)), size = 0.25) +
    #   geom_sf(data = tmp, aes(geometry = use_centroid, color = as.factor(RampID)), shape = 8) +
    #   ggtitle("Check dots are in centroid of uses in cells for each boat ramp") +
    #   theme(plot.title = element_text(hjust = 0.5)) +
    #   theme_classic() +
    #   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    #         legend.spacing.y = unit(0.75, "cm"))
    
    # making distance data frame
    dist <- as.data.frame(round(set_units(st_distance(tmp, BRtrips), km), 2)) # distance from use centroid to each BR
    names(dist) <- unique(BRtrips$RampID) 
    dist$gridID_vis <- tmp$gridID_vis 
    dist$RampID <- tmp$RampID 
    dist$id <- tmp$id 
    len <- length(dist) - 3
    dist %<>% gather("BRdist", "use.centroid_km.br", 1:len) 
    dist[] <- sapply(dist, as.numeric) # change all cols to numeric
    
    dist <- dist %>% 
      filter(RampID == BRdist) %>% 
      dplyr::select(-c(id, BRdist)) %>% 
      rename(gridID_alt = gridID_vis) # got to join to alt not visited
    
    dat <- as.data.frame(dat)
    
    dat <- left_join(dat, dist, by = c("RampID", "gridID_alt")) # join
    
    dat <- dat %>% 
      mutate(km_BR = ifelse(is.na(use.centroid_km.br), site.centroid_km.br,
                            use.centroid_km.br)) # distance tc to be measured from
    
    # validation plot
    ggplot(dat) +
      geom_sf(aes(geometry = asc_geom, fill = use.centroid_km.br), lwd = 0.25) +
      geom_sf(aes(geometry = br_geom), size = 3) +
      theme_classic() +
      scale_fill_distiller(palette = "Blues") +
      labs(fill = "km from BR", colour = "") +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
            legend.spacing.y = unit(0.75, "cm")) +
      facet_grid(.~fac_br_id) +
      ggtitle("Check distances look accurate")
```

## Choice and travel cost ($0.54/km)
```{r choice and travel cost}
    dat <- dat %>% 
      # mutate(centroidLong = st_coordinates(polyCentroid)[,1]) %>% # need this is
      # using dynamic variables
      # mutate(centroidLat = st_coordinates(polyCentroid)[,2]) %>%
      mutate(choice = ifelse(gridID_vis == gridID_alt, 1, 0)) %>% 
      mutate(travelCost = as.numeric(0.58*km_BR)) # 0.58 = ave fuel price/consumption


#'[#TODO:make a function which appends average for area fuel price to month fo visit for tc calculation ]  

    
    #'[#VALIDATE: TRUE, should equal number of trips]
    sum(dat$choice) == n_distinct(dat$TripID)
```

## Filtering SZ
  - Current sz will have no data and therefore no value so are removed
  - Simulated sz that have no data have no value so are  removed
```{r}
    # Remove sites with no recreational value
    dat <- dat %>%  filter(sz_value == 1 | is.na(sz_value))
    
    # remove observations that have visited the current sz (removed above)
    sz_current_w_use <- setdiff(unique(dat$gridID_vis), unique(dat$gridID_alt))
    
    dat2 <- dat %<>% filter(!(gridID_vis %in% sz_current_w_use))
    
    #'[#VALIDATE: TRUE, no difference between gridID_alt/vis as every cells needs data]
  
     setdiff(unique(dat$gridID_vis), unique(dat$gridID_alt))
     setdiff(unique(dat$gridID_alt), unique(dat$gridID_vis))
     
    
     
    # removing sz columns
    sz_to_remove <- dat %>% 
      dplyr::select(num_range("sz", 1:100)) %>% 
      summarise_all(sum) 
    
    sz_to_remove <- names(sz_to_remove[, which(sz_to_remove %in% 0)])
     
    dat_rum_sf <- dat %>% dplyr::select(-c(sz_to_remove))
    dat_rum_df <- dat %>% dplyr::select(-c(sz_to_remove, asc_geom, br_geom, centroid, dat_geom))
    dat_df <- dat %>% dplyr::select(-c(asc_geom, br_geom, centroid, dat_geom))
    
    #'[#TODO: check this]  
    nrow(dat) == n_distinct(dat_df$gridID_alt) * n_distinct(dat_df$TripID)
    
    
```

# Save
```{r save}
    # full grid including sites with no recreational value
    saveRDS(dat, paste0("./data/gpkg/sims/2.1_ascCS_",sim_name ,"_4283.gpkg"))
    write_csv(dat_df, paste0("./data/02_data/2.1_asc_", sim_name,".csv"))
    
    # full grid excluding sites with no recreational value
    saveRDS(dat_rum_sf, paste0("./data/gpkg/sims/2.1_ascCS_rum_", sim_name,"_4283.gpkg")) # script 3.1
    write_csv(dat_rum_df, paste0("./data/02_data/2.1_asc_rum_", sim_name,".csv")) # stata
    
    # simplified ascGrid for QGIS (geom, ID)
    st_write(gridID_vis,
             paste0("./data/gpkg/sims/2.1_ascCSQ", sim_name, "_4283.gpkg"),
             append = FALSE) # need to write with st_write to open in QGISRUMs
```
