# Wrangle data for ASC RUM

## Set up
```{r setup}
# knitr options
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, fig.align = 'center', fig.width = 10, fig.height = 7) 

# libraries 
library(tidyverse)
library(dplyr)
library(ggplot2)
library(sp)
library(raster)
library(rgeos)
library(rgdal)
library(sf)
library(lwgeom)
library(googledrive)
library(units)
library(nngeo)
library(magrittr)
library(todor)
library(smoothr)

# data
BR <- st_read("./data/gpkg/NingBR.gpkg") # boat ramps
BRtrips <- read.csv("./data/RAW/Trips_Ramp_out.csv") 
ntz <- st_read("./data/gpkg/ntz.gpkg") # current sz, state and commonwealth
grid_base <- st_read("./data/gpkg/mar_acBase.gpkg") # grid base
dat <- read_csv("data/01_data/1.2_dat.csv") # data

# functions
source("./functions/spatialFunc.R") # spatial functions
source("./functions/ascFunc.R")
source("./functions/theme.R")

sf_use_s2(FALSE)
```

```{r}
BRtrips <- read.csv("./data/02_data/Trips_Ramp_out.csv") %>% 
  filter(State_reg == "WA" & Name_reg == "West Coast") # filter to state and region

# check ramps names
# unique(BRtrips$Ramp_name)

BRtrips <- BRtrips %>% 
  filter(Ramp_name %in% c("Hillarys Marina", "Mindarie Marina", "Ocean Reef Marina", "Two Rocks Marina")) %>% 
  dplyr::select(Ramp_name, RampID, Month, TRIPS_ramp_prop_Grvt__5, upp_TRIPS_Grvt__5, low_TRIPS_Grvt__5) %>% 
  group_by(RampID, Ramp_name) %>%
  summarize(across(where(is.numeric), list(sum), .names = "{.col}")) %>% # summing across months to get annual estimates
  mutate(RampID = as.numeric(RampID))

BRtrips <- left_join(BRtrips, BR, by = c("RampID")) %>% # join to geometry
  st_as_sf()

write_csv(BRtrips, "./data/02_data/2.1_BRtrips.csv")
saveRDS(BRtrips, "./data/gpkg/2.1_BRtrips.gpkg")
```

## Simulation
```{r}
# sim_name <- "simog_split"
# sim <- st_read("./data/gpkg/simog_split.gpkg")

# sim_name <- "sim1_split"
# sim <- st_read("./data/gpkg/sim1_split.gpkg")

# sim_name <- "sim2_split"
# sim <- st_read("./data/gpkg/sim2_split.gpkg")

# sim_name <- "sim3_split"
# sim <- st_read("./data/gpkg/sim3_split.gpkg")

sim_name <- "sim4_split"
sim <- st_read("./data/gpkg/sim4_split.gpkg")

# sim %<>% mutate(id = row_number())
# 
# sim %<>% filter(id == 9)
# ggplot()+
#   geom_sf(data = sim, aes(fill = as.factor(id)))
```
## Data prep
All data filters have to be done before making the choice set.
```{r}
dat %<>% 
  filter(ActivityType == "Extractive") %>% # filtering activity
  filter(!is.na(UseLat), !is.na(UseLong)) %>% # filtering empty coords
  st_as_sf(coords = c("UseLong", "UseLat"), crs = 4283)  ## make sf object
  

dat %<>% st_crop(dat, grid_base) # removing outliers

# need one trip per row (longest site visited on trip)
dat %<>% 
  group_by(PersonID, TripID) %>% 
  slice(which.max(decDuration)) %>% 
  ungroup()

which(duplicated(dat$TripID) == TRUE) # Should be none
```

## Choice set
```{r message=FALSE}
grid <- ASCgrid(poly = grid_base, sz_current = npz, sz_sim = sim, point = dat, from = 0.03, by = 0.02, min_area = 0, crs = 4283, vert = FALSE) # crs using degrees
```
# simulation stats
```{r}
grid %>% group_by(sz, sz_current, sz_w_value) %>%  summarise("Area (km2)" = sum(area)) # area of sz with and without value
grid %>% group_by(sz) %>% summarise(use_count = sum(use_count)) # use count in and out sz

mp <- st_read("./data/gpkg/pmmp_4326.gpkg") %>% st_transform(crs(dat))
coast <- st_read("./data/gpkg/Coast_4283.gpkg") %>% st_transform(crs(mp))

sz_mp <- grid %>% filter(sz == 1) %>% st_union()

sz_area <- sum(as.numeric(round(set_units(st_area(sz_mp), km^2), 2))) # total area of sz
mp_area <- as.numeric(round(set_units(st_area(mp), km^2), 2)) # total area pf mp
per_sz <- round((sz_area/mp_area)*100, 2) # percentage of mp to be made sz

# mp$use_count <- lengths(st_intersects(mp, dat))

# calculate mp length of coast
# calculate length of coast that has sz 
# percentage of sz that have access to the shore

# mp_coast <- st_touches(mp, coast)
# peri <- st_perimeter(mp_coast)
# 
# ggplot()+
#   # geom_sf(data = mp) +
#   # geom_sf(data = sz_mp) +
#   geom_sf(data = mp_coast, colour = 'red')
```


## Attributes
```{r}
site_centroid <- st_centroid_within_poly(grid) # gets centroid of every grid cell
grid$site_centroid <- site_centroid$geom # appending centroid

# checkpoint
ggplot() +
  geom_sf(data = grid$geom) +
  geom_sf(data = grid$site_centroid, color = 'red')

# getting distance from each boat rampe to the centroid of the sites
BR <- st_transform(BR, crs(grid))
dist <- as.data.frame(set_units(st_distance(grid$site_centroid, BR), km)) 
names(dist) <- unique(BR$RampID)
gridID_alt <- grid$gridID_alt
dist$gridID_alt <- gridID_alt
dist <- dist %>% gather("BR", "site.centroid_km.BR", 1:length(dist) - 1) 
grid <- grid %>% left_join(dist, by = "gridID_alt") # join 

grid <- grid |> st_sf() |> # wont work without st_sf()
  dplyr::select(-site_centroid)
```

## Data wrangle
```{r}
# isolating grid id to append to dat
gridID_vis <- grid %>% 
  dplyr::select(gridID_alt, geom) %>% 
  distinct() 

# st_write(gridID_vis, 
#          paste0("./data/gpkg/2.1_acChoiceSetQ", sim_name, "_4283.gpkg"), 
#          append = FALSE) # need to write with st_write to open in QGISRUMs

dat <- st_crop(dat, grid) # removing outliers

# allocating sites to grids
dat  %<>% st_join(gridID_vis, left = T, join = st_intersects)

# selecting attributes, and tidying
dat %<>% 
  rename(gridID_vis = gridID_alt) %>% 
  dplyr::select(TripID, PersonID, gridID_vis, Site) %>% 
  distinct() %>% 
  mutate(BR = ifelse(Site %in% "Hillarys", "104", NA),
         BR = ifelse(Site %in% "Mindarie", "105", BR),
         BR = ifelse(Site %in% "OceanReef", "106", BR),
         BR = ifelse(Site %in% "TwoRocks", "112", BR)) %>% 
  # mutate(UseLong = st_coordinates(.)[,1]) %>% 
  # mutate(UseLat = st_coordinates(.)[,2]) %>% 
  dplyr::select(-Site) %>% 
  as.data.frame()

# checkpoint
which(is.na(dat$gridID_alt) == TRUE) # should be none
```

## Join data to choice set
```{r join}
og_dat <- dat # copying original data test join

dat <- inner_join(dat, grid, by = "BR") # FALSE

# checkpoint
a <- as.data.frame(table(og_dat$BR)) %>% rename(launches = Freq)
b <- as.data.frame(table(grid$BR)) %>%  rename(n_sites = Freq)
c <- left_join(a, b)
c$times <- c$launches*c$n_sites
test = sum(c$times)
test == nrow(dat) # this should be TRUE
```

## Distance to BR
Distance from the boat ra,mp to the centroid of data points in each grid that launched from that boat ramp - if not takes centroid of grid cell. 
```{r message=FALSE}
dat <- dat %>% 
  st_as_sf() # turn to sf object

# get centroid of use points
tmp <- dat %>%
  group_by(gridID_vis, BR) %>%
  summarise(use_centroid = st_union(geometry)) %>%
  st_centroid() %>%
  ungroup() %>%
  mutate(id = row_number())

# making distance data frame
dist <- as.data.frame(set_units(st_distance(tmp, BR), km)) # distance from use centroid to each BR
names(dist) <- unique(BR$RampID) 
dist$gridID_vis <- tmp$gridID_vis 
dist$BR <- tmp$BR 
dist$id <- tmp$id 
len <- length(dist) - 3 
dist <- dist %>% gather("BRdist", "use.centroid_km.BR", 1:len) # gather
dist$use.centroid_km.BR <- as.numeric(dist$use.centroid_km.BR) 

dist <- dist %>% 
  filter(BR == BRdist) %>% 
  dplyr::select(gridID_vis, BR, use.centroid_km.BR) %>% 
  rename(gridID_alt = gridID_vis) # got to join to alt not visited

dat <- as.data.frame(dat)

dat <- left_join(dat, dist, by = c("BR", "gridID_alt")) # join

dat <- dat %>% 
  mutate(site.centroid_km.BR = as.numeric(site.centroid_km.BR)) %>% 
  mutate(km_BR = ifelse(is.na(use.centroid_km.BR), site.centroid_km.BR, use.centroid_km.BR)) # distance tc to be measured from
```

## Choice and travel cost ($0.54/km)
```{r choice and travel cost}
dat <- dat %>% 
  # mutate(centroidLong = st_coordinates(polyCentroid)[,1]) %>% # need this is using dynamic variables
  # mutate(centroidLat = st_coordinates(polyCentroid)[,2]) %>%
  mutate(choice = ifelse(gridID_vis == gridID_alt, 1, 0)) %>% 
  mutate(travelCost = as.numeric(0.58*km_BR)) %>%  # 0.58 = ave fuel price/consumption
  dplyr::select(-c(geometry, site.centroid_km.BR, use.centroid_km.BR))
```

## Filtering SZ
  - Current sz will have no data and therefore no value so are removed
  - Simulated sz that have no data have no value so are  removed
```{r}
# remove all current sz (sz_curret is made from gridID_alt)
dat %<>% filter(sz_current != 1)

# remove observations that have visited the current sz (removed above)
sz_current_w_use <- setdiff(unique(dat$gridID_vis), unique(dat$gridID_alt))
# length(sz_current_w_use) # number of current sz that have use in it
# length(which(dat$gridID_vis %in% sz_current_w_use))/length(unique(dat$gridID_alt)) # n trips to sz current

dat %<>% filter(!(gridID_vis %in% sz_current_w_use))

dat %<>% filter(sz_wo_value != 1)

# test: should be no difference between gridID_alt and gridID_vis because all grids ahould have at least 1 data point.
 setdiff(unique(dat$gridID_vis), unique(dat$gridID_alt))
 setdiff(unique(dat$gridID_alt), unique(dat$gridID_vis))

 #removing sz columns
sz_to_remove <- dat %>% 
  dplyr::select(num_range("sz", 1:100)) %>% 
  summarise_all(sum) 

sz_to_remove <- names(sz_to_remove[, which(sz_to_remove %in% 0)])
 
dat_rum_sf <- dat %>% dplyr::select(-c(sz_to_remove))
dat_rum_df <- dat %>% dplyr::select(-c(sz_to_remove, sz_current, sz_wo_value, geom))

# plots
sz_w_value_plot <- ggplot() +
  geom_sf(data = grid$geom, aes(fill = as.factor(grid$sz_w_value)), lwd = 0.25) +
  xlab("Longitude") + ylab("Latitude") +
  labs(fill = "Value") +
#   geom_sf(data = dat$geom, size = 0.5) +
  my_theme +
  ggtitle("SZ with value")

sz_wo_value_plot <- ggplot() +
  geom_sf(data = grid$geom, aes(fill = as.factor(grid$sz_wo_value)), lwd = 0.25) +
  xlab("Longitude") + ylab("Latitude") +
  labs(fill = "Value") +
#   geom_sf(data = dat, size = 0.5) +
  my_theme +
  ggtitle("SZ without value")

sz_w_value_plot
sz_wo_value_plot
```

## Save outputs
```{r save}

sim_name <- "test"
# plot
ggsave(sz_w_value_plot, path = "./plots/ASC RUM", filename = paste("acSZwValue_", sim_name, ".png", sep = ''), width = 4, height = 4)

ggsave(sz_wo_value_plot, path = "./plots/ASC RUM", filename = paste("acSZwoValue_", sim_name, ".png", sep = ''), width = 4, height = 4)

# data with grids cells with no value
saveRDS(dat, paste0("./data/gpkg/2.1_acChoiceSet_",sim_name ,"4283.gpkg"))
write_csv(dat, paste0("./data/02_data/2.1_ac", sim_name,".csv"))

# data without grids cells with no value, for model
saveRDS(dat_rum_sf, paste0("./data/gpkg/2.1_acChoiceSet_rum", sim_name,"_4283.gpkg")) # script 3.1
write_csv(dat_rum_df, paste0("./data/02_data/2.1_ac", sim_name,"_rum.csv")) # stata
```