# Setup
```{r setup, include = FALSE}
rm(list = ls())

knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, cache = TRUE, fig.align = 'center', fig.width = 10, fig.height = 7) 

# libraries
library(tidyverse)
library(readxl)
library(magrittr)
library(MASS)
library(matrixStats)
library(dplyr)
library(data.table)
library(gtable)
library(gridExtra)
library(sp)
library(raster)
library(rgeos)
library(rgdal)
library(sf)
library(readxl)

# data
# dat <- readRDS("./data/gpkg/2.1_ChoiceSet_4283.gpkg")  # this is the generalised data with static site attributes
# # dat <- readRDS("./data/gpkg/2.1_ChoiceSet_4283.gpkg")  # this is the generalised data with static site attributes
# BRtrips <- readRDS("./data/gpkg/2.1_BRtrips.gpkg") # This is produced by the "TripAllocation.R" script (Matt)
dat <- readRDS("./data/gpkg/sims/2.1_dat.gpkg")

# Plotting
# mp <- st_read("./data/gpkg/pmmp_4326.gpkg")
# # mp_bbox <- st_read("./data/gpkg/pscmpBbox_4326.gpkg")
# NPZ <- st_read("./data/gpkg/ntz_4283.gpkg")

grid <- dat %>% as.data.frame() %>%
  distinct(gridID_alt, .keep_all = TRUE) %>%
  dplyr::select(gridID_alt, geometry) %>%
  st_as_sf()
```

Model Data
Multiple model
b: Data with coefficients from model using trip data
v: covariance matrix
```{r}
# base model
b <- read_excel("./data/03_data/attmod_base.xlsx") %>% dplyr::select(Vars, Coef)

a <- nrow(b) + 2
v <- read_excel("./data/03_data/attmod_base.xlsx", sheet = "v", col_names = rep("x", a)) %>%
  dplyr::select(-x...1 ) %>%
  distinct()
```
# Rearrange data
Aim is to have matrix with same row names and column names
```{r rearrange, include = FALSE}
    colnames(v) <- c("vars", v$x...2)
    v %<>% dplyr::select(-vars)
    vars <- b$Vars #use
    
nrow(b) == length(v) # should be true
length(v) == nrow(v) # should be true

# write.csv(b, "./data/03_data/bioec_coefs.csv")
# write.csv(v, "./data/03_data/bioec_covmat.csv")
```

# Transformation
Need to transform data in choice ste to match transformations used in model
```{r transformations, include = FALSE}
dat <- dat %>%
  # mutate(depthsquared = Depth*Depth) %>%
   mutate(arealog = log(as.numeric(area) + 1))
```

# Error 
Generates coefficients drawn from within the ci of each variable.
```{r error, include = FALSE}
# number of samples (all estimates)
    n1 <- 100

# draws n1 numbers (nrow) from within a multivariate normal distribution for each variable (length)
# mu = vector of means, Sigma = covarience matrix
    mvn_b <- mvrnorm(n1, mu = b$Coef, Sigma = as.matrix(v)) 
# dim(mvn_b)[1] == n1 # should be true
# dim(mvn_b)[2] == nrow(b) # should be true
```

# Utility
Calculate the utility (Vj) for every alternative, each individual observation in dat represents an alternative. 

We expect recreator  (i) to choose the site (j) with the highest expected utility. The utility  (Vj) of a site is a function of the observed attributes (B) of those sites and associated error (e).

      Vj = B(depth, travel cost etc.) + e 

```{r utility, include = FALSE}

  bid <- unique(b$Vars) 
  id_vars <- dat %>% dplyr::select(matches(bid))
  vars <- c(names(id_vars), "isl_adj") #TODO automate
      
    # turn defined vars into matrix from dat, make sure there is no geometry
        rum_matrix <- as.matrix(dat[, vars])
        
    # create an empty utility matrix
        Vj <- matrix(NA, nrow = nrow(rum_matrix), ncol = n1) 
    
# calculate fixed utility for every alternative (nrows), n1 times from multivariate random normal distribution. %*% is a matrix multiplication where...
# We are multiplying the values in the rum_matrix (real data) by the error coefficients matrix (mvn_b)? so we have 100 utility values for each variable -? why? 

    for (i in 1:n1) { 
      Vj[ ,i] <- as.vector(rum_matrix %*% as.matrix(mvn_b[i,]))  
    }
    
    Vj <- exp(Vj) # get exponential of utility
```

The probability of a recreator (i) choosing site (j) is the exponential of the the utility of the site (j) divided by the sum of the utility of all the potential alternatives the recreator could have choosen (including site j). This is a conditional logit formula. 

    probj = exp(Vj)/sum(exp(Vj))
    
```{r Probabilities, include = FALSE}
# Probabilities

    Vj <- as.data.table(Vj) # make data.table to save RAM
    cols <- colnames(Vj) # save the names of the draws for later
    Vj$br <- as.integer(dat$Site) # run function by BR
    Vj$ID <- as.integer(dat$gridID_alt) # add gridID_alt for group_by
    Vj$tc <- dat$tc3L # add travel cost

# Calculate probability of visit and logsums: pre-simulation
# probability function:  probj = exp(Vj)/sum(exp(Vj))

    f1 <- function(x) if (is.integer(x)) {x} else {x/sum(x)} 
    
    # applying probability function to every utility cell in Vj by boat ramp

    # npz <- dat %>%  filter(NPZ == 1) # remove for 5 km grid
    # 
    # Vj[Vj$ID %in% npz$gridID_alt, cols] <- 0 # remove for 5 km grid
    p <- Vj[ , lapply(.SD, f1), by = br, .SDcols = c(cols, "ID")] 
    p <- p[order(br, ID)]
    dat %<>% arrange(Site, gridID_alt)
    table(dat$gridID_alt == p$ID) # this should all be true (they do now)
    
    # alternatives presented to the person - sum of exp(Vj)
     logsum_base <- Vj[ , lapply(.SD, sum), by = br, .SDcols = cols]
    
    p <- p[ , br := NULL]
    p <- p[ , ID := NULL]
    p <- as.matrix(p)   

# Calculates probability of visit summary statistics by row. This is the base probability of each site, for each trip and the error. 
# For each site p has 100 estimates of the p of visiting that site me give me tyhe average, median, ci..

    dat$p.mean_base <- rowMeans(p) # median
    dat$p.low_base <- rowQuantiles(p, probs = sqrt(0.025)) # lower ci
    dat$p.upp_base <- rowQuantiles(p, probs = 1 - sqrt(0.025)) # upper ci 
    
# save output
    # write.csv(dat, "./ignore/03_data/3.1_basedat.csv")
    # write.csv(logsum_base, "./ignore/03_data/3.1_logsum_base.csv")

```

# Sim 1
```{r closures, include = FALSE}
# Assign closures
# 5km grid
# sim <- st_read("./data/gpkg/sim1_5km.gpkg")
# sim %<>% rename(gridID_alt = fid)
# ntz <- c(sim$gridID_alt)

# 1km grid
sim <- st_read("./data/gpkg/Sim 1.gpkg")
ntz <- sim$gridID_alt

# Calculate probability of visit and logsums: post-simulation
    
    # Vj[Vj$ID %in% npz$gridID_alt, cols] <- 0 # remove for 5km grid
    Vj[Vj$ID %in% ntz, cols] <- 0
    p <- Vj[ , lapply(.SD, f1), by = br, .SDcols = c(cols, "ID")]
    p <- p[order(br, ID)]
    dat %<>% arrange(BR, gridID_alt)
    table(dat$gridID_alt == p$ID) # should all be true
    
    logsum_sim <- Vj[ , lapply(.SD, sum), by = br, .SDcols = cols] #18876 trips
    
    p <- p[ , br := NULL]
    p <- p[ , ID := NULL]
    p <- as.matrix(p)
    
    
# Calculates probability of visit summary statistics by row. This is the simulated probability of each site, for each trip and the error. 

    dat$p.mean_sim <- rowMeans(p)
    dat$p.low_sim <- rowQuantiles(p, probs = sqrt(0.025)) #saving outputs back to p
    dat$p.upp_sim <- rowQuantiles(p, probs = 1 - sqrt(0.025)) #Note below about weird probs
    
    # ggplot() +
    #   geom_sf(data = sites) +
    #   geom_sf(data = sim, fill = "red")
```

# Extract Welfare
```{r welfare, include = FALSE}
# Extracting the welfare impact
    W <- logsum_sim[, -1] -  logsum_base[, -1]
    W <- t(W)/mvn_b[ ,1] # length of this does not match n1/mvn_b
    W <- t(W)

# Summaries across draws
    W <- as.matrix(W)
    logsum_sim$w.mean <- rowMeans(W)
    logsum_sim$w.low <- rowQuantiles(W, probs = sqrt(0.025)) 
    logsum_sim$w.upp <- rowQuantiles(W, probs = 1 - sqrt(0.025))
    logsum_sim$upp.ci <- logsum_sim$w.upp - logsum_sim$w.mean
    
    dat %<>%  mutate(BR = as.integer(BR))
  
    dat %<>% left_join(logsum_sim[ , c("br", "w.mean", "w.upp", "w.low", "upp.ci")], by = c("BR" = "br"))

    # # Welfare impact of closure
    W_impact_trip <- logsum_sim %>%
      summarise("Annual lower welfare impact per trip ($)" = round(sum(w.low), 2),
                "Annual mean welfare impact per trip ($)" = round(sum(w.mean), 2),
                "Annual upper welfare impact per trip ($)" = round(sum(w.upp), 2))

    W_impact_trip
    
    # png(paste0("./plots/Attribute RUM/", mod,".png"), height = 50*nrow(W_impact_trip), width =
    #       200*ncol(W_impact_trip))
    # grid.table(W_impact_trip)
    # dev.off()

# save output
  # write.csv(dat, "./ignore/03_data/3.1_Wsimdat.csv")
  # write.csv(logsum_sim, "./ignore/03_data/3.1_logsum_sim.csv")
```
# Annual welfare impacts per boat ramp

## Annual BR visits
This chunk sums monthly br visits and gravity functions to get annual estimates and joins to rum data. 
  
```{r brtrip data prep, include = FALSE}
# checking brs match
    setdiff(BRtrips$RampID, dat$BR)
    setdiff(dat$BR, BRtrips$RampID)

# joining trip data to dat
    RUM <- dat %>% left_join(.,BRtrips, by = c("BR" = "RampID" ), suffix = c("", "_Ramp"))
```

## Multipling welfare impacts across annual br visits

This chunk merges estimates of trip numbers at each ramp each annually (BRtrips) with estimates of
the probability of visiting each grid (by annually) given a particular boat ramp was selected (dat). 

  - Number of annual trips per br * the W impact of each br
  
```{r br multiplication, include = FALSE}
    RUM_W <- RUM %>% distinct(.,BR, .keep_all = TRUE) %>%
      mutate(across(starts_with("TRIPS_ramp_prop"), ~.x*w.mean, .names = "welf_mean_{.col}" ),
             across(starts_with("upp_TRIPS"), ~.x*w.upp, .names = "welf_upp_{.col}" ),
             across(starts_with("low_TRIPS"), ~.x*w.low, .names = "welf_low_{.col}" ))

# check match BR in shp file
    setdiff(RUM_W$BR, BRtrips$RampID)
    setdiff(BRtrips$RampID, RUM_W$BR)
    
# join to shp file
    # RUM_W %<>% left_join(BR[,c("RampID")], by = c("BR" = "RampID"), suffix = c("","_ramp"))
    
# multiplied br numbers by corresponding probabilities. Note that because we adjusted the probs above to 
# sqrt(0.025) we should now have accurate 95% CIs. 
    
    RUM %<>% mutate(across(starts_with("TRIPS_ramp_prop"), ~.x*p.mean_sim, .names = "mean_sim_{.col}" ),
                             across(starts_with("TRIPS_ramp_prop"), ~.x*p.mean_base, .names =
                                      "mean_base_{.col}"),
                             across(starts_with("upp_TRIPS"), ~.x*p.upp_sim, .names = "upp_sim_{.col}" ),
                             across(starts_with("upp_TRIPS"), ~.x*p.upp_base, .names = "upp_base_{.col}" ),
                             across(starts_with("low_TRIPS"), ~.x*p.low_sim, .names = "low_sim_{.col}" ),
                             across(starts_with("low_TRIPS"), ~.x*p.low_base, .names = "low_base_{.col}" ))

# Sum by grid (over ramps) to get use at grid level
      RUM_grid <- RUM %>% group_by(gridID_alt) %>%  # was just grouping by ID
      summarise(across(starts_with(c("mean_sim_","upp_sim_","low_sim_","mean_base_", "upp_base_", 
                                     "low_base_")), sum))
# join it to geometry
    RUM_grid <- right_join(grid, RUM_grid, by = c("gridID_alt"))
```

# Welfare impacts
```{r totW, include = FALSE}
W_impact <- RUM_W %>% summarise(round(across(starts_with(c("welf_low", "welf_mean","welf_upp")), sum)))

W_impact %<>%  rename("Annual lower welfare impact ($)" = welf_low_low_TRIPS_Grvt__5,
                      "Annual mean welfare impact ($)" = welf_mean_TRIPS_ramp_prop_Grvt__5,
                      "Annual upper welfare impact ($)" = welf_upp_upp_TRIPS_Grvt__5)

W_impact

# png("./plots/Attribute RUM/att_s1.W_impact.png", height = 50*nrow(W_impact), width = 200*ncol(W_impact))
# grid.table(W_impact)
# dev.off()
```



```{r sc_spimpact, include = FALSE}
# add planning area and crop to it
mp_bbox <- st_read("./data/gpkg/StudySite_4283.gpkg") %>% st_transform(crs(RUM_grid))
mpbr <- as.data.frame(st_intersects(BRtrips, mp_bbox, sparse = FALSE))
BRtrips$mpbr <- mpbr$V1
sp_mpbr <- BRtrips %>% filter(mpbr == TRUE)
mpbr <- sp_mpbr$RampID

# bbox <- RUM_grid %>% filter(BR %in% mpbr)
# 
RUM_grid %<>%
     mutate(mean_diff = mean_sim_TRIPS_ramp_prop_Grvt__5 - mean_base_TRIPS_ramp_prop_Grvt__5,
            upp_diff = upp_sim_upp_TRIPS_Grvt__5 - upp_base_upp_TRIPS_Grvt__5,
            low_diff = low_sim_low_TRIPS_Grvt__5 - low_base_low_TRIPS_Grvt__5) %>% 
  filter(mean_diff >= 0)
# 
# bbox <- RUM_grid %>% st_crop(mp_bbox)
# mp %<>% st_transform(crs(bbox))
# mp_crop <- mp %>% st_crop(bbox)
# mp_comm <- NPZ %>% st_crop(bbox)

presim <- ggplot(RUM_grid) +
      xlab("Longitude") + ylab("Latitude") +
      geom_sf(aes(fill = mean_base_TRIPS_ramp_prop_Grvt__5), lwd = 0.05) +
      scale_fill_distiller(palette = "YlGnBu", direction = 1) +
      labs(fill = "Trips\nannually") +
  my_theme +
  ggtitle(" Pre-simulation ") +
  geom_sf(data = mp, alpha = 0, lwd = 0.25) +
  geom_sf(data = NPZ, lwd = 0.25) +
  geom_sf(data = sp_mpbr, color = 'red', fill = 'red', size = 0.5)

ggsave(path = "./plots/Attribute RUM", filename = paste("att_s1.presim.png", sep = ''), plot = presim, width = 8, height = 4)

postsim <- ggplot(RUM_grid) +
      xlab("Longitude") + ylab("Latitude") +
      geom_sf(aes(fill = mean_sim_TRIPS_ramp_prop_Grvt__5), lwd = 0.05) +
      scale_fill_distiller(palette = "YlGnBu", direction = 1) +
      labs(fill = "Trips\nannually") +
  geom_sf(data = sim, fill = 'black') +
  my_theme +
   ggtitle(" Post-simulation ") +
  geom_sf(data = mp, alpha = 0, lwd = 0.25) +
  geom_sf(data = NPZ, lwd = 0.25) +
  geom_sf(data = sp_mpbr, color = 'red', fill = 'red', size = 0.5)

ggsave(path = "./plots/Attribute RUM", filename = paste("att_s1.postsim.png", sep = ''), width = 8, height = 4)

diff <- ggplot(RUM_grid) +
      xlab("Longitude") + ylab("Latitude") +
      geom_sf(aes(fill = mean_diff), lwd = 0.05) +
      scale_fill_distiller(palette = "YlGnBu", direction = -1) +
      labs(fill = "Change in\nannual trips") +
      my_theme +
    geom_sf(data = mp, alpha = 0, lwd = 0.25) +
    geom_sf(data = NPZ, lwd = 0.25) +
  geom_sf(data = sp_mpbr, color = 'red', fill = 'red', size = 0.5) +
  geom_sf(data = sim, lwd = 0.05, fill = "black")

ggsave(path = "./plots/Attribute RUM", filename = paste("att_s1.diff.png", sep = ''), width = 8, height = 4)

presim
postsim
diff
```

## Welfare impacts per boat ramp
```{r welfare_bar, include = FALSE}
RUM_W$BR <- as.factor(RUM_W$BR)

W_BRimpact <- RUM_W %>% group_by(BR, Ramp_name) %>%
      summarise(round(across(starts_with(c("welf_low", "welf_mean","welf_upp")), sum))) %>% 
  filter(welf_mean_TRIPS_ramp_prop_Grvt__5 > 0)

W_BRimpact_bar <- W_BRimpact %>%
  ggplot(.) +
  geom_bar(aes(x = BR, y = welf_mean_TRIPS_ramp_prop_Grvt__5), stat = "identity",  fill = col) +
  geom_errorbar(aes(x = BR, ymin = welf_low_low_TRIPS_Grvt__5, ymax = welf_upp_upp_TRIPS_Grvt__5), width = 0.2) +
  my_theme +
  scale_x_discrete(drop = T, labels = W_BRimpact$Ramp_name) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  ylab("Annual welfare impacts on\nrecreational fishers ($)") +
  xlab("Boat Ramp")

W_BRimpact_bar
ggsave(path = "./plots", filename = paste("sim1.W_BRimpact_bar.png", sep = ''), plot = W_BRimpact_bar, width = 8, height = 4)


W_BRimpact %<>%  rename("Annual lower welfare impact ($)" = welf_low_low_TRIPS_Grvt__5,
                        "Annual mean welfare impact ($)" = welf_mean_TRIPS_ramp_prop_Grvt__5,
                        "Annual upper welfare impact ($)" = welf_upp_upp_TRIPS_Grvt__5)

W_BRimpact
png("./plots/sim1.W_BRimpact.png", height = 50*nrow(W_BRimpact), width = 150*ncol(W_BRimpact))
grid.table(W_BRimpact)
dev.off()
```

```{r}
# for (i in W_BRimpact$BR) {
#   br <- BRtrips %>% filter(RampID %in% i)
# bbox <- RUM_grid %>%  filter(BR %in% i)
# mp_crop <- mp %>% st_crop(bbox)
# mp_comm <- NPZ %>% st_crop(bbox)
# 
# presim <- RUM_grid %>% filter(BR == i) %>%
#     ggplot(.) +
#       xlab("Longitude") + ylab("Latitude") +
#       geom_sf(aes(fill = mean_base_TRIPS_ramp_prop_Grvt__5), lwd = 0.05) +
#       scale_fill_distiller(palette = "YlGnBu", direction = 1) +
#       labs(fill = "Trips\nannually") +
#       my_theme +
#     geom_sf(data = mp_crop, alpha = 0, lwd = 0.25) +
#   geom_sf(data = mp_comm, alpha = 0, lwd = 0.25) +
#   ggtitle(paste(br$Ramp_name, " pre-simulation ")) +
#   geom_sf(data = br, color = 'red', fill = 'red', size = 1)
# 
# ggsave(path = "./plots", filename = paste("sim1.", br$Ramp_name,"_presim.png", sep = ''), plot = presim, width = 8, height = 4)
# print(presim)
# 
# sim <- RUM_grid %>% filter(BR == i & gridID_alt %in% ntz)
# 
# postsim <- RUM_grid %>% filter(BR == i) %>%
# ggplot(.) +
#       xlab("Longitude") + ylab("Latitude") +
#       geom_sf(aes(fill = mean_sim_TRIPS_ramp_prop_Grvt__5), lwd = 0.05) +
#       scale_fill_distiller(palette = "YlGnBu", direction = 1) +
#       labs(fill = "Trips\nannually") +
#   geom_sf(data = mp_crop, alpha = 0, lwd = 0.25) +
#   geom_sf(data = mp_comm, alpha = 0, lwd = 0.25) +
#   geom_sf(data = br, color = 'red', fill = 'red', size = 1) +
#   geom_sf(data = sim, fill = "black", lwd = 0.05) +
#   ggtitle(paste(br$Ramp_name, " post-simulation")) +
#   my_theme
# 
# ggsave(path = "./plots", filename = paste("sim1.", br$Ramp_name,"_postsim.png", sep = ''), plot = postsim, width = 8, height = 4)
# print(postsim)
# }
```

```{r difference plots}
# RUM_grid %<>%
#      mutate(mean_diff = mean_sim_TRIPS_ramp_prop_Grvt__5 - mean_base_TRIPS_ramp_prop_Grvt__5,
#             upp_diff = upp_sim_upp_TRIPS_Grvt__5 - upp_base_upp_TRIPS_Grvt__5,
#             low_diff = low_sim_low_TRIPS_Grvt__5 - low_base_low_TRIPS_Grvt__5)
# 
# for (i in W_BRimpact$BR) {
#   br <- BRtrips %>% filter(RampID %in% i)
#   bbox <- RUM_grid %>%  filter(BR %in% i)
#   mp_crop <- mp %>% st_crop(bbox)
#   mp_comm <- NPZ %>% st_crop(bbox)
#   sim <- RUM_grid %>% filter(BR == i & gridID_alt %in% ntz)
#   
# diff <- RUM_grid %>% filter(BR == i) %>%
#     ggplot(.) +
#       xlab("Longitude") + ylab("Latitude") +
#       geom_sf(aes(fill = mean_diff), lwd = 0.05) +
#       scale_fill_distiller(palette = "YlGnBu", direction = -1) +
#       labs(fill = "Change in\nannual trips") +
#       my_theme +
#     geom_sf(data = mp_crop, alpha = 0, lwd = 0.25) +
#     geom_sf(data = mp_comm, alpha = 0, lwd = 0.25) +
#       ggtitle(paste(br$Ramp_name)) +
#   geom_sf(data = sim, fill = "black", lwd = 0.05) +
#     geom_sf(data = br, color = 'red', fill = 'red', size = 1)
# 
# ggsave(path = "./plots", filename = paste("sim1.", br$Ramp_name,"_diff.png", sep = ''), plot = diff, width = 8, height = 4)
# print(diff)
# }
```