```{r response stats}
Approached <- dat %>%
  distinct(PersonID, .keep_all = TRUE) 
nApproached <- nrow(Approached) # the number of people approached

nNE <- length(which(Approached$Agreement == "NE")) # number of people approached that were ineligible to participate in the survey
nAgree <- length(which(Approached$Agreement == "Yes")) # number of people approached who agreed to participate in the survey, including repeats 
nRefusals <- length(which(Approached$Agreement == "No")) # number of people approached who refused to take part in the survey, including repeats
nIndPeople <- length(which(Approached$Agreement %in% "Yes" & Approached$PrevInter %in% "No")) # the number of individual people surveyed
nMultiAgree <- nAgree - nIndPeople # number of multi-agreement surveys (total number of surveys repeated by previous participants - excluding their first survey)

nSpPoints <- nrow(dat) - nRefusals - nNE # the total number of spatial data points
nMultiUse <- nSpPoints - nAgree # the number of spatial point that were part of multi-use trips (excluding first instance)

RefusalRate <- round((nRefusals/(nApproached - nNE))*100) # refusal rate
ResponseRate <- round((nAgree/(nApproached - nNE))*100) # response rate
```
<br><br>

> MEG Mission statement: **Optimising human use of the ocean**

***
# Data {.tabset}
Data for recreational marine use in daterance was collected through face-to-face surveys at boat launches.  The surveys facilitated the collection of multi-use data and multi-greement data.The survey had a **`r ResponseRate`% response rate** with **`r nAgree` participants**, of those **`r nIndPeople` survey were first instance or single agreement surveys** and **`r nMultiAgree` were multi-agreement surveys** (survey participant surveyed for a second or more times). This produced a **total of `r nSpPoints` spatial use data points**, **`r nMultiUse` of which were part of multi-use trips** (excluding first instance).

```{r data prep, include = FALSE}
yes <- dat %>% 
  filter(Agreement == "Yes") # agreement data (filtering out refusals and non-eligibles)

if (boat == TRUE) { # boat based (extractive and non-extractive)
  actb <- yes %>% filter(SiteType == "Boat")
  # ifelse(nrow(actb) > 1, eval_actb <- TRUE, eval_actb <- FALSE)
} 

if (exists("actb")) {
  ifelse(nrow(actb) > 1, eval_actb <- TRUE, eval_actb <- FALSE)
} else if (!exists("actb")) {
  eval_actb <- FALSE
}


if (shore == TRUE) { # shore based (extractive and non-extractive)
  acts <- yes %>% filter(SiteType == "Shore") 
} 

if (exists("acts")) {
  ifelse(nrow(acts) > 1, eval_acts <- TRUE, eval_acts <- FALSE)
} else if (!exists("acts")) {
  eval_acts <- FALSE
}

if (extractive == TRUE) { # extractive - boat and shore
  ex <- yes %>% filter(Activity == "Fishing") 
} 

if (exists("ex")) {
  ifelse(nrow(ex) > 1, eval_ex <- TRUE, eval_ex <- FALSE)
} else if (!exists("ex")) {
  eval_ex <- FALSE
}

if (boat == TRUE & extractive == TRUE) { # extractive - boat  
  exb <- yes %>% filter(SiteType == "Boat" & ActivityType %in% c("Extractive", "Both")) 
} 

if (exists("exb")) {
  ifelse(nrow(exb) > 1, eval_exb <- TRUE, eval_exb <- FALSE)
} else if (!exists("exb")) {
  eval_exb <- FALSE
}

if (shore == TRUE & extractive == TRUE) {  # extractive - shore  
  exs <- yes %>% filter(SiteType == "Shore" & ActivityType %in% c("Extractive", "Both"))
} 

if (exists("exs")) {
  ifelse(nrow(exs) > 1, eval_exs <- TRUE, eval_exs <- FALSE)
} else if (!exists("exs")) {
  eval_exs <- FALSE
}

if (nonextractive == TRUE) { # non-extractive - boat and shore 
  nex <- yes %>% filter(Activity != "Fishing") 
} 

if (exists("nex")) {
  ifelse(nrow(nex) > 1, eval_nex <- TRUE, eval_nex <- FALSE)
} else if (!exists("nex")) {
  eval_nex <- FALSE
}

if (boat == TRUE & nonextractive == TRUE) {  # non-extractive - boat 
  nexb <- yes %>% filter(SiteType == "Boat" & ActivityType %in% c("NonExtractive"))
}

if (exists("nexb")) {
  ifelse(nrow(nexb) > 1, eval_nexb <- TRUE, eval_nexb <- FALSE)
} else if (!exists("nexb")) {
  eval_nexb <- FALSE
}

if (shore == TRUE & nonextractive == TRUE) {  # non-extractive - shore  
nexs <- yes %>% filter(SiteType == "Shore" & ActivityType %in% c("NonExtractive"))
}

if (exists("nexs")) {
  ifelse(nrow(nexs) > 1, eval_nexs <- TRUE, eval_nexs <- FALSE)
} else if (!exists("nexs")) {
  eval_nexs <- FALSE
}
```


## Response
The plot below shows the number of agreements, refusals and non-eligible people that were approached in the sampling process. 
```{r bar_response}
dat %>% 
  distinct(PersonID, .keep_all = TRUE) %>% 
  ggplot() +
   aes(x = fct_infreq(Agreement)) + 
   geom_bar(fill = col1) +
   labs(x = "Agreement to participate", y = paste("Frequency (n = ", nrow(Approached), ")", sep = "")) +
   geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme
ggsave("./plots/explore/agree.png", width = 8, height = 4) 
```


```{r bar_sitetype, eval = boat & shore}
# text
asis_output("## Site type\nThe plot below shows the number of boat and shore based activities that were captured in the data.")

# plot
yes %>% 
  ggplot() +
   aes(x = fct_infreq(SiteType)) +
   geom_bar(fill = col1) +
   labs(x = "Site Type", y = paste("Frequency (n = ", nrow(yes), ")", sep = "")) +
   geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme
 ggsave("./plots/explore/bar_sitetype.png", width = 8, height = 4) 
```


```{r bar_year}
# heading
if (length(unique(yes$facYear)) > 1) {
  asis_output("## Year\nThe plot below shows the how the data is distributed across years.")
}

# plot
if (length(unique(yes$facYear)) > 1) {
year<- yes %>%
  ggplot()+
  aes(x = as.factor(facYear)) +
  geom_bar(fill = col1) +
  labs(x = "Year", y = paste("Number of Participants (n = ", nrow(yes), ")", sep = "")) +
  geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme
 ggsave("./plots/explore/bar_year.png", width = 8, height = 4) 
 year
}

```

# Demographics  {.tabset}
***

## Locality
The plot below shows how many of the participants were residents and visitors.
```{r bar_locality}
ggplot(yes) + 
   aes(x = fct_infreq(Resident)) +
   geom_bar(fill = col1) +
   labs(x = "Resident", y = paste("Frequency (n = ", nrow(yes), ")", sep = "")) +
   geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme
 ggsave("./plots/explore/bar_locality.png", width = 8, height = 4) 
```

## Age
The plot below shows the ages of participants.
```{r bar_age}
ggplot(yes) + 
   aes(x = binAge) +
   geom_bar(fill = col1) +
   labs(x = "Age of Participant", y = paste("Frequency (n = ", nrow(yes), ")", sep = "")) +
   geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme
 ggsave("./plots/explore/bar_age.png", width = 8, height = 4) 
```

## Gender
The plot below shows the gender of participants.
```{r bar_gender}
ggplot(yes) + 
   aes(x = fct_infreq(Sex)) +
   geom_bar(fill = col1) +
   labs(x = "Sex of Participant", y = paste("Frequency (n = ", nrow(yes), ")", sep = "")) +
   geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme
 ggsave("./plots/explore/bar_gender.png", width = 8, height = 4) 
```


```{r bar_exavid, eval = eval_ex}
# text
asis_output("## Fishing Avidity\nThe plot below shows the fishing avidity of participants.")

# plot
ggplot(yes) + 
   aes(x = bin_exTimes12m) +
   geom_bar(fill = col1) +
   labs(x = "Number of times fished in last 12 months", y = paste("Frequency (n = ", nrow(yes), ")", sep = "")) +
   geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme
 ggsave("./plots/explore/bar_exavid.png", width = 8, height = 4) 
```

# Use
***

```{r kdens_all}
# text
if (boat == TRUE & shore == FALSE) {
  asis_output("## Overall {.tabset}\nThis map shows the kernel density of all boat based uses, extractive and non-extractive.")
} else if (boat == FALSE & shore == TRUE) {
  asis_output("## Overall {.tabset}\nThis map shows the kernel density of all shore based uses, extractive and non-extractive.")
} else if (boat == TRUE & shore == TRUE) {
  asis_output("## Overall {.tabset}\nThe maps below show the kernel density of boat and shore based recreational usesm extractive and non-extractive.") 
}

# plot
if (boat == TRUE & shore == FALSE) {
  a <- actb %>% 
    ggplot() +
    stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), geom =
                      "polygon", bins = 10) +
    scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
    scale_alpha(guide = 'none') +
    geom_sf(data = coast, lwd = 0.07) +
    map_theme +
    labs(title = "Boat based activities", x = "", y = "")
  ggsave("./plots/explore/kdens_actb.png", width = 8, height = 4) 
  print(a)
} else if (boat == FALSE & shore == TRUE) {
  b <- acts %>%
    ggplot() +
    stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), geom =
                      "polygon", bins = 10) +
    scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
    scale_alpha(guide = 'none') +
    geom_sf(data = coast, lwd = 0.07) +
    map_theme +
  labs(title = "Shore based activities")
  ggsave("./plots/explore/kdens_acts.png", width = 8, height = 4) 
  print(b)
} else if (boat == TRUE & shore == TRUE & landscape == TRUE) {
  a <- actb %>% 
    ggplot() +
    stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), geom =
                      "polygon", bins = 10) +
    scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
    scale_alpha(guide = 'none') +
    geom_sf(data = coast, lwd = 0.07) +
    map_theme +
    labs(title = "Boat based activities", x = "", y = "")
  ggsave("./plots/explore/kdens_actb.png", width = 8, height = 4) 
  
  b <- acts %>%
    ggplot() +
    stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), geom =
                      "polygon", bins = 10) +
    scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
    scale_alpha(guide = 'none') +
    geom_sf(data = coast, lwd = 0.07) +
    map_theme +
  labs(title = "Shore based activities")
  ggsave("./plots/explore/kdens_acts.png", width = 8, height = 4) 
  print(a)
  print(b)
} else if (boat == TRUE & shore == TRUE & portrait == TRUE) {
  a <- actb %>% 
    ggplot() +
    stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), geom =
                      "polygon", bins = 10) +
    scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
    scale_alpha(guide = 'none') +
    geom_sf(data = coast, lwd = 0.07) +
    map_theme +
    labs(title = "Boat based activities", x = "", y = "")
  ggsave("./plots/explore/kdens_actb.png", width = 8, height = 4) 
  
  b <- acts %>%
    ggplot() +
    stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), geom =
                      "polygon", bins = 10) +
    scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
    scale_alpha(guide = 'none') +
    geom_sf(data = coast, lwd = 0.07) +
    map_theme +
  labs(title = "Shore based activities")
  ggsave("./plots/explore/kdens_acts.png", width = 8, height = 4) 
  print(a + b) 
}
```


```{r bar_sp_acttype, eval = extractive & nonextractive}
# text
asis_output("### Activity Type\nThe plots below shows how many surveys are relevant to extractive and non-extractive activities and there spatial distribution.")

# plots
if (landscape == TRUE) {
  a <- ggplot(yes) +
    aes(x = fct_infreq(ActivityType)) +
    geom_bar(fill = col1) +
    labs(x = "Activity Type", y = paste("Frequency (n = ", nrow(yes), ")", sep = "")) +
    geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) + 
    my_theme
  ggsave("./plots/explore/bar_acttype.png", width = 8, height = 4) 
 
 b <- basemap + 
  geom_sf(data = spdat, aes(color = ActivityType), size = 0.5) +
  scale_color_manual(values = c(col1, col2), name = "")
  map_theme
  ggsave("./plots/explore/sp_acttype.png", width = 8, height = 4) 
print(a)
print(b)
} else if (portrait == TRUE) {
  a <- ggplot(yes) +
    aes(x = fct_infreq(ActivityType)) +
    geom_bar(fill = col1) +
    labs(x = "Activity Type", y = paste("Frequency (n = ", nrow(yes), ")", sep = "")) +
    geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
    my_theme
  ggsave("./plots/explore/bar_acttype.png", width = 8, height = 4) 

 b <- basemap +
  geom_sf(data = spdat, aes(color = ActivityType), size = 0.5) +
  scale_color_manual(values = c(col1, col2), name = "") +
  map_theme
ggsave("./plots/explore/sp_acttype.png", width = 8, height = 4) 
a + b
ggsave("./plots/explore/bar_sp_acttype.png", width = 8, height = 4)
print(a + b)
}
```



```{r bar_actfreq, eval = extractive & nonextractive}
# text
asis_output("### Activities\nThe plots below shows what activities participants conducted and there spatial distribution.")

# plots
if (landscape == TRUE) {
  a <- yes %>%  # replace with next line if map is portrait
    ggplot(aes(y = fct_rev(fct_infreq(Activity)))) +
    geom_bar(stat="count", fill = col1) +
    labs(x = paste("Frequency (n = ", nrow(yes), ")", sep = ""), 
         y = paste("Activity (n = ",   length(na.omit(unique(yes$Activity))), ")", sep="")) +
    geom_text(stat='count', aes(label=..count..), hjust = -0.5, size = 2.5) +
    my_theme
  ggsave("./plots/explore/bar_actfreq.png", width = 8, height = 4)
  
  b <- basemap +
    geom_sf(data = spdat, aes(color = Activity), size = 0.7) +
    scale_color_brewer(palette = "Set3", name ="") +
    map_theme
  ggsave("./plots/explore/sp_actfreq.png", width = 8, height = 4)
  
  print(a)
  print(b)
  
} else if (portrait == TRUE) {
  a <- yes %>%
    ggplot(aes(y = fct_rev(fct_infreq(Activity)))) +
    geom_bar(stat="count", fill = col1) +
    labs(x = paste("Frequency (n = ", nrow(yes), ")", sep = ""), 
         y = paste("Activity (n = ",   length(na.omit(unique(yes$Activity))), ")", sep="")) +
    geom_text(stat='count', aes(label=..count..), hjust = -0.5, size = 2.5) +
    my_theme
  ggsave("./plots/explore/bar_actfreq.png", width = 8, height = 4)
  
  b <- basemap +
    geom_sf(data = spdat, aes(color = Activity), size = 0.7) +
    scale_color_brewer(palette = "Set3", name ="") +
    map_theme
  ggsave("./plots/explore/sp_actfreq.png", width = 8, height = 4)
  
  a + b
  ggsave("./plots/explore/bar_sp_acttype.png", width = 8, height = 4)
  print(a + b)
}
```


```{r ex_heading, eval = extractive}
asis_output("## Extractive use\n***") # header
```


```{r bar_exb, eval = eval_exb}
# text
asis_output("### Boat based {.tabset}\n")
asis_output("#### Fishing type\n")

# plot
exb %>%
 ggplot(aes(x = fct_infreq(FishingType))) +
  geom_bar(stat='count', position = position_dodge2(reverse = TRUE), fill = col1) +
  labs(x = paste("Fishing Type (n = ", length(na.omit(unique(exb$FishingType))), ")", sep=""), y = paste("Frequency (n = ", nrow(exb), ")", sep = "")) +
  geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme
ggsave("./plots/explore/bar_exb.png", width = 8, height = 4)
```


```{r sp_exb, eval = eval_exb}
# text
asis_output("#### Spatial distribution and kernel density\n")

# sf of boat based fishing for kernel density plot - keep lat/long independent of geometry
sp_exb <- exb %>% st_as_sf(coords = c("UseLong", "UseLat"), crs = 4283) %>% 
  mutate(UseLong = sf::st_coordinates(.)[,1], UseLat = sf::st_coordinates(.)[,2])

# plot
if (landscape == TRUE) {
  a <- basemap +
    geom_sf(data = sp_exb, aes(colour = FishingType), size = 0.5) +
    scale_color_brewer(palette = "Set3") +
    labs(color='Boat based fishing') +
    map_theme
  ggsave("./plots/explore/sp_exb.png", width = 8, height = 4)

  b <- sp_exb %>%
    ggplot() +
    stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), geom = "polygon", bins = 10) +
    scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
    scale_alpha(guide = 'none') +
    geom_sf(data = coast, lwd = 0.07) + 
    map_theme +
    labs(x = "", y = "")
  ggsave("./plots/explore/kdens_exb.png", width = 8, height = 4)
  
  print(a)
  print(b)
} else if (portrait == TRUE) {
  a <- basemap +
    geom_sf(data = sp_exb, aes(colour = FishingType), size = 0.5) +
    scale_color_brewer(palette = "Set3") +
    labs(color='Boat based fishing') +
    map_theme
  ggsave("./plots/explore/sp_exb.png", width = 8, height = 4)

  b <- sp_exb %>%
    ggplot() +
    stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), geom = "polygon", bins = 10) +
    scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
    scale_alpha(guide = 'none') +
    geom_sf(data = coast, lwd = 0.07) + 
    map_theme +
    labs(x = "", y = "")
  ggsave("./plots/explore/kdens_exb.png", width = 8, height = 4)
  
  a + b 
  ggsave("./plots/explore/sp_kdens_exb.png", width = 8, height = 4)
  print(a + b)
}
```


```{r bar_exs, eval = eval_exs}
# text
asis_output("### Shore based {.tabset}\n")
asis_output("#### Fishing type\n")

# plot
exs %>%
  ggplot(aes(x = fct_infreq(FishingType))) +
  geom_bar(stat='count', position = position_dodge2(reverse = TRUE), fill = col1) +
  labs(x = paste("Fishing Type (n = ", length(na.omit(unique(exs$FishingType))), ")", sep=""), y = paste("Frequency (n = ", nrow(exs), ")", sep = "")) +
  geom_text(stat='count', aes(label=..count..), vjust = -0.5, size = 2.5) +
  my_theme
ggsave("./plots/explore/bar_exs.png", width = 8, height = 4)
```


```{r sp_exs, eval = eval_exs}
# text
asis_output("#### Spatial distribution and kernel density\n")

# sf of boat based fishing for kernel density plot - keep lat/long independent of geometry
sp_exs <- st_as_sf(exs, coords = c("UseLong", "UseLat"), crs = 4283) %>% 
  mutate(UseLong = sf::st_coordinates(.)[,1], UseLat = sf::st_coordinates(.)[,2])

# plot
if (landscape == TRUE) {
  a <- basemap +
    geom_sf(data = sp_exs, aes(colour = FishingType), size = 0.5) +
    scale_color_brewer(palette = "Set3") +
    labs(color='Boat based fishing') +
    map_theme
  ggsave("./plots/explore/sp_exb.png", width = 8, height = 4)

  b <- sp_exs %>%
    ggplot() +
    stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), geom = "polygon", bins = 10) +
    scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
    scale_alpha(guide = 'none') +
    geom_sf(data = coast, lwd = 0.07) + 
    map_theme +
    labs(x = "", y = "")
  ggsave("./plots/explore/kdens_exb.png", width = 8, height = 4)
  
  print(a)
  print(b)
} else if (portrait == TRUE) {
  a <- basemap +
    geom_sf(data = sp_exs, aes(colour = FishingType), size = 0.5) +
    scale_color_brewer(palette = "Set3") +
    labs(color='Boat based fishing') +
    map_theme
  ggsave("./plots/explore/sp_exb.png", width = 8, height = 4)

  b <- sp_exs %>%
    ggplot() +
    stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), geom = "polygon", bins = 10) +
    scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
    scale_alpha(guide = 'none') +
    geom_sf(data = coast, lwd = 0.07) + 
    map_theme +
    labs(x = "", y = "")
  ggsave("./plots/explore/kdens_exb.png", width = 8, height = 4)
  
  a + b 
  ggsave("./plots/explore/sp_kdens_exb.png", width = 8, height = 4)
  print(a + b)
}
```

```{r nex_heading, eval = nonextractive}
asis_output("## Non-extractive use\n***") # header
```


```{r bar_nexb, eval = eval_nexb}
# text
asis_output("### Boat based {.tabset}\n")
asis_output("#### Activity\n")

# plot
nexb %>%
  ggplot(aes(y = fct_rev(fct_infreq(Activity)))) +
  geom_bar(stat="count", fill = col1) +
  labs(x = paste("Frequency (n = ", nrow(nexb), ")", sep = ""), y = paste("Activity (n =", length(na.omit(unique(nexb$Activity))), ")", sep="")) +
  geom_text(stat='count', aes(label=..count..), hjust = -0.5, size = 2.5) +
  my_theme
ggsave("./plots/explore/bar_nexb.png", width = 8, height = 4)
```


```{r sp_nexb, eval = eval_nexb}
asis_output("#### Spatial distribution and kernel density\n")

# sf of boat based fishing for kernel density plot - keep lat/long independent of geometry
sp_nexb <- st_as_sf(nexb, coords = c("UseLong", "UseLat"), crs = 4283) %>% 
  mutate(UseLong = sf::st_coordinates(.)[,1], UseLat = sf::st_coordinates(.)[,2])

# plot
if (landscape == TRUE) {
  a <- basemap +
    geom_sf(data = sp_nexb, aes(color = Activity), size = 1) +
    scale_color_brewer(palette = "Set3") +
    map_theme
  ggsave("./plots/explore/sp_nexb.png", width = 8, height = 4)
  
  b <- sp_nexb %>%
    ggplot() +
    stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), 
                    geom = "polygon", bins = 10) +
    scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
    scale_alpha(guide = 'none') +
    geom_sf(data = coast, lwd = 0.07) +
    my_theme + theme(axis.text.x = element_text(angle = 90))
  ggsave("./plots/explore/kdens_nexb.png", width = 8, height = 4)
  
  print(a)
  print(b)
} else if (portrait == TRUE) {
  a <- basemap +
    geom_sf(data = sp_nexb, aes(color = Activity), size = 1) +
    scale_color_brewer(palette = "Set3") +
    map_theme
  ggsave("./plots/explore/sp_nexb.png", width = 8, height = 4)
  
  b <- sp_nexb %>%
    ggplot() +
    stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), 
                    geom = "polygon", bins = 10) +
    scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
    scale_alpha(guide = 'none') +
    geom_sf(data = coast, lwd = 0.07) +
    my_theme + theme(axis.text.x = element_text(angle = 90))
  ggsave("./plots/explore/kdens_nexb.png", width = 8, height = 4)
  
  a + b 
  ggsave("./plots/explore/sp_kdens_nexb.png", width = 8, height = 4)
  print(a + b)
}
```


```{r bar_nexs, eval = eval_nexs}
# text
asis_output("### Shore based {.tabset}\n")
asis_output("#### Activity\n")

# plot
nexs %>%
  ggplot(aes(y = fct_rev(fct_infreq(Activity)))) +
  geom_bar(stat="count", fill = col1) +
  labs(x = paste("Frequency (n = ", nrow(nexs), ")", sep = ""), 
       y = paste("Activity (n =", length(na.omit(unique(nexs$Activity))), ")", sep="")) +
  geom_text(stat='count', aes(label=..count..), hjust = -0.5, size = 2.5) +
  my_theme
ggsave("./plots/explore/bar_nexs.png", width = 8, height = 4)
```


```{r sp_nexs, eval = eval_nexs}
asis_output("#### Spatial distribution and kernel density\n")

# sf of boat based fishing for kernel density plot - keep lat/long independent of geometry
sp_nexs <- st_as_sf(nexs, coords = c("UseLong", "UseLat"), crs = 4283) %>%
  mutate(UseLong = sf::st_coordinates(.)[,1], UseLat = sf::st_coordinates(.)[,2])

# plot
if (landscape == TRUE) {
  a <- basemap +
    geom_sf(data = sp_nexs, aes(color = Activity), size = 1) +
    scale_color_brewer(palette = "Set3") +
    map_theme
  ggsave("./plots/explore/sp_nexs.png", width = 8, height = 4)
  
  b <- nexs %>%
    ggplot() +
    stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), 
                    geom = "polygon", bins = 10) +
    scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
    scale_alpha(guide = 'none') +
    geom_sf(data = coast, lwd = 0.07) +
    map_theme
  ggsave("./plots/explore/kdens_nexs.png", width = 8, height = 4)
  
  print(a)
  print(b)
} else if (portrait == TRUE) {
  a <- basemap +
    geom_sf(data = sp_nexs, aes(color = Activity), size = 1) +
    scale_color_brewer(palette = "Set3") +
    map_theme
  ggsave("./plots/explore/sp_nexs.png", width = 8, height = 4)
  
  b <- nexs %>%
    ggplot() +
    stat_density_2d(aes(x = UseLong, y = UseLat, fill = stat(level), alpha = ..level..), 
                    geom = "polygon", bins = 10) +
    scale_fill_distiller(palette = "YlGnBu", direction = -1, name = "Kernal\nDensity") +
    scale_alpha(guide = 'none') +
    geom_sf(data = coast, lwd = 0.07) +
    map_theme
  ggsave("./plots/explore/kdens_nexs.png", width = 8, height = 4)
  
  a + b 
  ggsave("./plots/explore/sp_kdens_nexs.png", width = 8, height = 4)
  print(a + b)
}
```