nlogit posteestimation
- calculates choice probabilities 
- calcaluates logsums 

# Setup
```{r setup, include = FALSE}
  rm(list = ls())

knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, cache = TRUE, fig.align = 'center', fig.width = 10, fig.height = 7) 

# libraries
library(tidyverse)
library(readxl)
library(magrittr)
library(MASS)
library(matrixStats)
library(dplyr)
library(data.table)
library(gtable)
library(gridExtra)
library(sp)
library(raster)
library(rgeos)
library(rgdal)
library(sf)

# simulation 
model <- "nex" #'[#NOTE: choose model] 
sim <- "sim4" #'[#NOTE: choose simulation] 
cost <- "fcflt_spgam"  #'[#NOTE: choose cost variable] # not needed for attatasc

b <- read_excel(paste0("./data/03_data/", model, ".xlsx")) # coefs for model (marginal utilities)
a <- nrow(b) + 2
v <- read_excel(paste0("./data/03_data/", model, ".xlsx"), sheet = "v", col_names = rep("x", a)) %>% distinct()

# NEX RUMN paper
dat <- read_csv(paste0("./data/02_data/2.1_nex_", sim, ".csv")) # NEX
fishers <- read_csv(paste0("./data/02_data/2.1_nex_ex_", sim, ".csv")) # comprable EX for NEX
```

# Wrangle data
dat: contains a row for  (nrow(dat)) = every available option (n grid cells, excluding current sz and sim sz with no recreational value) * per trip (n trips). The choice column indicated what site the recreator actually chose. 
```{r rearrange, include = FALSE}
    b %<>% rename(cov = ...1) %>% 
      mutate(cov = ifelse(Vars %in% "age", paste0(cov, "_", Vars), cov)) %>% 
      mutate(cov = ifelse(!(Vars %in% c("_cons", "age")), Vars, cov)) %>% 
      dplyr::select(-Vars)
  
    v %<>% 
      # mutate(x...1 = ifelse(x...2 %in% "travelcost", "tc", x...1)) %>%  
      mutate(x...1 = ifelse(x...2 %in% "age", paste0(x...1, "_", x...2), x...1)) %>% 
      mutate(x...1 = ifelse(!(x...2 %in% c("_cons", "age")), x...2, x...1)) %>%  
      dplyr::select(-x...2) %>% 
      distinct()

    colnames(v) <- c("vars", v$x...1)
    v %<>% dplyr::select(-vars)
    vars <- b$cov #use

    #'[#VALIDATE: TRUE/TRUE/check min and max, extreme values will skew model (issue 2.1?)]
    nrow(b) == length(v) # should be true
    length(v) == nrow(v) # should be true
    summary(b$Coef) # check min and max, the more extreme the values the more inflated or deflated the results will be. Most likely a problem with script 2.1 or model.
    
    tau <- b[86:89, ]
    tau %<>% mutate(cov = substr(cov, 1, 3)) 
    tau %<>% rename(cset = cov, tau = Coef) 
    dat %<>% left_join(tau)
    
    b_asc <- b[c(1, 5:85),]
    b_nest <- b[2:4,]
    
    v_asc <- v[c(1, 5:85),]
    v_asc <- v_asc %>% dplyr::select(-contains("age") & -contains("tau"))
    
    v_nest <- v[2:4,]
    v_nest <- v_nest %>% dplyr::select(contains("age"))
```

# Utility matrix
# Bottom level alternatives
```{r}
## Generate error from cov-var matrix
    n1 <- 1000 # number of samples (all estimates)
    mvn_b <- mvrnorm(n1, mu = b_asc$Coef, Sigma = as.matrix(v_asc))  # generate error
    dim(mvn_b)[1] == n1 # should be true
    
## Make rum matirx which alings with b cov
    b_asc %<>% mutate(cov = ifelse(!str_detect(cov, "fc"), paste0("gridid_", cov), cov)) # make names the same
    bid <- unique(b_asc$cov) 
    rum_matrix <- as.matrix(dat[, bid]) # extract data 
      
## Calc utility
    vj_bl <- matrix(NA, nrow = nrow(rum_matrix), ncol = n1) # create an empty utility matrix
        
    for (i in 1:n1) { 
        vj_bl[ ,i] <- as.vector(rum_matrix %*% as.matrix(mvn_b[i,]))  
        }
        
    which(is.infinite(as.matrix(vj_bl)))
    
## Conditional Probability
   tau <- as.vector(dat$tau)
   condp_num <- exp(vj_bl/tau)
   
    dat %<>%
        mutate(gridid_int = group_indices(., gridid_alt),
               tripid_int = group_indices(., trip_id),
               cset_int = group_indices(., cset))  # making id integers
    
   condp_num <- as.data.table(condp_num) # make data.table to save RAM
   condp_num_cols <- colnames(condp_num) # save the names of the draws for later
   condp_num$trip <- as.integer(dat$tripid_int) # add 
   condp_num$cset <- as.integer(dat$cset_int) # add 
   
   condp_num %<>%  relocate(cset, .before = 1)
   condp_num %<>%  relocate(trip, .before = 1)
   
  condp_dom <- condp_num %>% 
    group_by(trip, cset) %>% 
    summarise(across(V1:V1000, sum)) %>% 
    rename_with(~ gsub("V", "dom_", .), starts_with("V"))
    
  condp_dom %<>% left_join(condp_num)
  condp_dom %<>% dplyr::select(1:1002) 
  
  condp_num <- as.matrix(condp_num[,3:1002])
  condp_dom <- as.matrix(condp_dom[,3:1002])

  condp <- condp_num/condp_dom
  
  dat$lci_condp <- rowQuantiles(condp, probs = sqrt(0.025)) # lower ci
  dat$m_condp <- rowQuantiles(condp, probs = sqrt(0.5)) # lower ci
  dat$uci_condp <- rowQuantiles(condp, probs = 1 - sqrt(0.025)) # upper ci
  
  iv <- log(condp_dom)
  
# P1: Pr that agent chooses nest
## utility of nest
    n1 <- 1000 # number of samples (all estimates)
    mvn_b <- mvrnorm(n1, mu = b_nest$Coef, Sigma = as.matrix(v_nest))  # generate error
    dim(mvn_b)[1] == n1 # should be true
    
## Make rum matirx which alings with b cov
  
    rum_matrix <- as.matrix(dat[, c("age", "age", "age")]) # extract data 
      
## Calc utility
    vj_k <- matrix(NA, nrow = nrow(rum_matrix), ncol = n1) # create an empty utility matrix
        
    for (i in 1:n1) { 
        vj_k[ ,i] <- as.vector(rum_matrix %*% as.matrix(mvn_b[i,]))  
        }
        
    which(is.infinite(as.matrix(vj_k)))
  
  iv <- iv*tau
  
  p1_num <- exp(vj_k+iv)
  
  ## I AM HERE
  p1_num$trip <- dat$tripid_int
  
  p1_dom <- p1_num %>% 
    group_by(trip) %>% 
    summarise(across(V1:V1000, sum)) %>% 
    rename_with(~ gsub("V", "dom_", .), starts_with("V"))
    
  condp_dom %<>% left_join(condp_num)
  condp_dom %<>% dplyr::select(1:1002) 
  
  
  

# p1 denominator
tmp <- dat %>% group_by(trip_id) %>% distinct(lci_p1_num) %>% summarise(lci_p1_dom = sum(lci_p1_num))
dat %<>% left_join(tmp) 

tmp <- dat %>% group_by(trip_id) %>% distinct(p1_num) %>% summarise(p1_dom = sum(p1_num))
dat %<>% left_join(tmp) 

tmp <- dat %>% group_by(trip_id) %>% distinct(uci_p1_num) %>% summarise(uci_p1_dom = sum(uci_p1_num))
dat %<>% left_join(tmp) 

any(dat$lci_p1_dom > dat$p1_dom) # FALSE
any(dat$uci_p1_dom < dat$p1_dom) # FALSE

# p1
dat %<>%
  mutate(lci_p1 = lci_p1_num/lci_p1_dom,
         p1 = p1_num/p1_dom,
         uci_p1 = uci_p1_num/uci_p1_dom)

any(dat$lci_p1 > dat$p1) # TRUE
any(dat$uci_p < dat$p1) # FALSE
  
  
  
  

   
    dat[which(dat$trip_id %in% 259), c("gridid_alt", "cset", "trip_id" ,"lci_condp", "m_condp", "uci_condp")]
   
    
   f1 <- function(x) if (is.integer(x)) {x} else {x/tau}
    
   condp <- Vj[ , lapply(.SD, f1), .SDcols = c(cols)] # 
    
    
      # dat %<>%
      #   mutate(gridid_int = group_indices(., gridid_alt),
      #    tripid_int = group_indices(., trip_id))  # making id integers
    #     
    # Vj <- as.data.table(Vj) # make data.table to save RAM
    # cols <- colnames(Vj) # save the names of the draws for later
    # # Vj$tau <- dat$tau
    # Vj$ID <- as.integer(dat$gridid_int) # run function by BR
    # # Vj$tc <- dat$fcflt_spgam # add travel cost
    # # Vj$trip_id <- as.integer(dat$trip_id) # add trip id
    # Vj$br <- as.integer(dat$ramp_id) # add trip id
    
    
    # nx <- length(Vj) - n1
    # nx <- length(Vj) - nx
    #     
    # condp_Vj <- Vj %>%  mutate(across(1:nx, ~ . / tau)) # divide utilities by tau - NESTED LOGIT ONLY
    # # Vj %<>%  mutate(across(1:nx, ~ . / tau)) # divide utilities by tau - NESTED LOGIT ONLY
    # 
    # ## FOR ALL
    #  condp_Vj %<>% mutate(across(1:nx, exp)) # get exponential of utility
     # Vj %<>% mutate(across(1:nx, exp)) # get exponential of utility
    
        # Vj <- exp(Vj) # get exponential of utility
        
        #'[#VALIDATE: TRUE, exp(Vj) should make all values positive]
       # identical(which(Vj < 0), integer(0))
       which(is.infinite(as.matrix(Vj)))
     
     # identical(which(condp_Vj < 0), integer(0))
     #  which(is.infinite(as.matrix(condp_Vj)))   
    
       
       
         for (i in unique(dat$cset)) {
    cset = ifelse(dat$cset == i, 1, 0)
    dat[,length(dat) + 1] <- cset
    colnames(dat)[ncol(dat)] <- paste0(i, "_tau")
  }
       
       
      dat %<>% mutate(b_n_age = age,
                      s_e_age = age,
                      s_n_age = age) 
    
```


# Error 
Generates coefficients drawn from within the ci of each variable. Draws n1 numbers (nrow) from within a multivariate normal distribution for each variable (length); mu = vector of means, Sigma = covarience matrix
```{r error, include = FALSE}
    n1 <- 1000 # number of samples (all estimates)
    mvn_b <- mvrnorm(n1, mu = b$Coef, Sigma = as.matrix(v)) 

# checkpoint
dim(mvn_b)[1] == n1 # should be true
dim(mvn_b)[2] == nrow(b) # should be true
```

# Utility matrix

The coef of the logit model gives us the marginal utilities. This is the extra utility an agent receives for one extra unit, given all the other parameters stay constant. 

Calculate the utility (Vj) for every alternative, each individual observation in dat represents an alternative. 

We expect recreator  (n) to choose the site (j) with the highest expected utility. The utility  (Vj) of a site is a function of the observed attributes (B) of those sites and associated error (e).

      Vnj = Bj + Ej

      Vj = B(depth + travel cost etc.) + e 
      
The errors are assumed to follow the generalized extreme value (GEV) distribution type 1 (Gumbel distribution). This distribution is a normal distribution with a slight larger tail on one side. The errors are also assumed to be identical and independently distributed (i.i.d. assumption). This means there cannot be any correlation in the errors. Therefore the site definitions have to have no correlated errors, and therefore have to be different enough that the error don't correlated. If the researcher thinks there is some unobserved correlation in the errors, then the alternatives can be grouped into nests. For example boat fishing and shore fishing and inherently different, therefore you could cluster shore and boat based sites into a boat and shore nest. Doing a nested model relaxes the iid assumption and allows for some correlation within nests, but no correlation between nests. 

```{r utility, include = FALSE}
# ASC NEX RUM
      b %<>% mutate(cov = ifelse(!str_detect(cov, "fc|age|tau"), paste0("gridid_", cov), cov)) # NEX 
      bid <- unique(b$cov) 
 
      dat %<>% mutate(b_n_age = age,
                      s_e_age = age,
                      s_n_age = age) 
      
  for (i in unique(dat$cset)) {
    cset = ifelse(dat$cset == i, 1, 0)
    dat[,length(dat) + 1] <- cset
    colnames(dat)[ncol(dat)] <- paste0(i, "_tau")
  }

      rum_matrix <- as.matrix(dat[, bid])
      
    
    # create an empty utility matrix
        Vj <- matrix(NA, nrow = nrow(rum_matrix), ncol = n1)
        
# calculate observed utility for every alternative (nrows), n1 times from multivariate random normal distribution. We are multiplying the values in the rum_matrix (real data) by the error coefficients matrix (mvn_b)? so we have n1 utility values for each variable. 
    
# The loop below often fails: non-confromable. This means that you need to check the names in the vectors above to make sure the are the same as the model and mach the data set in the right order
        
        for (i in 1:n1) { 
          Vj[ ,i] <- as.vector(rum_matrix %*% as.matrix(mvn_b[i,]))  
        }
        
      # dat %<>%
      #   mutate(gridid_int = group_indices(., gridid_alt),
      #    tripid_int = group_indices(., trip_id))  # making id integers
    #     
    # Vj <- as.data.table(Vj) # make data.table to save RAM
    # cols <- colnames(Vj) # save the names of the draws for later
    # # Vj$tau <- dat$tau
    # Vj$ID <- as.integer(dat$gridid_int) # run function by BR
    # # Vj$tc <- dat$fcflt_spgam # add travel cost
    # # Vj$trip_id <- as.integer(dat$trip_id) # add trip id
    # Vj$br <- as.integer(dat$ramp_id) # add trip id
    
    
    # nx <- length(Vj) - n1
    # nx <- length(Vj) - nx
    #     
    # condp_Vj <- Vj %>%  mutate(across(1:nx, ~ . / tau)) # divide utilities by tau - NESTED LOGIT ONLY
    # # Vj %<>%  mutate(across(1:nx, ~ . / tau)) # divide utilities by tau - NESTED LOGIT ONLY
    # 
    # ## FOR ALL
    #  condp_Vj %<>% mutate(across(1:nx, exp)) # get exponential of utility
     # Vj %<>% mutate(across(1:nx, exp)) # get exponential of utility
    
        # Vj <- exp(Vj) # get exponential of utility
        
        #'[#VALIDATE: TRUE, exp(Vj) should make all values positive]
       # identical(which(Vj < 0), integer(0))
       which(is.infinite(as.matrix(Vj)))
     
     # identical(which(condp_Vj < 0), integer(0))
     #  which(is.infinite(as.matrix(condp_Vj)))
    
       # data for testing manually   
       # t <- dat %>% filter(trip_id %in% 259)
       # t <- t[, c("trip_id", "choice", "fcflt_spgam", "cset", "gridid_alt", "tau")]
       # t %<>% arrange(gridid_alt)
       # write.csv(t, "./data/03_data/test.csv")
```

# Conditional probability matrix
```{r}

  dat %<>%
        mutate(gridid_int = group_indices(., gridid_alt),
         tripid_int = group_indices(., trip_id))  # making id integers

    Vj <- as.data.table(Vj) # make data.table to save RAM
    cols <- colnames(Vj) # save the names of the draws for later
    Vj$ID <- as.integer(dat$gridid_int) # add gridID_alt for group_by
    Vj$tc <- dat$fcflt_spgam # add travel cost
    Vj$tau <- dat$tau

## condp scalar
dat %<>%
  mutate(lci_condp_scale = lci_tot_vj/lci_tau,
         condp_scale = tot_vj/tau, 
         uci_condp_scale = uci_tot_vj/uci_tau)

    f1 <- function(x) if (is.integer(x)) {x} else {x/tau}
    
    condp <- Vj[ , lapply(.SD, f1), .SDcols = c(cols)] # applying probability function to every cell in Vj - no groupd when BOAT AND SHORE BASED should be same length as data
    
    
      
    # Calculate probability of visit and logsums: pre-simulation
    f1 <- function(x) if (is.integer(x)) {x} else {x/sum(x)}




    # logsum_condp <- condp_Vj[ , lapply(.SD, sum), by = trip_id, .SDcols = cols]
    # 
    # logsum_condp <- log(logsum_condp[, -1])
    # w <- as.matrix(logsum_condp)
    # logsum_condp$mean_logsum <- rowQuantiles(w, probs = sqrt(0.025)) # lower ci
    
```


## 3 values 
 If you manged to find out how to extract the correct confiende intervals  so you only have 3 values. 

### Wrangle
The data is in the form of the model output from stata and is formatted like a table. 
```{r rearrange, include = FALSE}
# b %<>%
#   rename(cov = "fchoice",
#          coef = "Coefficient",
#          stdp = "Std. err.",
#          z_stat = "z",
#          p_value = "P>|z|",
#          lci = "[95% conf. interval]",
#          uci = "...7") # rename model output table
# 
# b %<>% mutate(cov = ifelse(cov == "_cons", lag(cov), cov)) # filling gaps in cov with correct gridid
# b %<>% filter(rowSums(is.na(.)) <= 2) # filter rows you dont need
# b %<>% mutate(cov = ifelse(str_detect(cov, "tau"), substr(cov, 1, 3), cov)) # change name of taus to match dat
# 
# asc <- b[2:82,] # isolate ASCs: one less that you would expect because base is missing
# dat %<>% left_join(asc, by = c("gridid_alt" = "cov")) # append ASCs
# dat %<>% mutate_at(vars(coef, lci, uci), ~ ifelse(is.na(.), 0, .))
# 
# tau <- b[83:86,] # isolate taus
# tau %<>%
#   dplyr::select(-c(3:5)) %>%
#   rename(cset = cov, tau = coef, lci_tau = lci, uci_tau = uci)
# dat %<>% left_join(tau, by = "cset")  # append taus
# 
# # b$coef[1] <- -3
# dat %<>% mutate(lci_tc = b$lci[1],
#                 tc = b$coef[1],
#                 uci_tc = b$uci[1]) # appending tc coef to data
# 
# # dat[which(dat$trip_id %in% 259), c("cset", "gridid_alt", "coef", "lci", "uci", "tau", "lci_tau", "uci_tau", "tc", "lci_tc", "uci_tc")]
```

# Base
```{r}
# calculate utility of tc
dat %<>%
  mutate(lci_vj_tc = lci_tc * fcflt_spgam,
         vj_tc = tc * fcflt_spgam,
         uci_vj_tc = uci_tc * fcflt_spgam)

any(dat$lci_vj_tc > dat$vj_tc) # FALSE
any(dat$uci_vj_tc < dat$vj_tc) # FALSE

# calculate total utility
dat %<>%
  mutate(lci_tot_vj = lci_vj_tc + coef,
         tot_vj = vj_tc + coef,
         uci_tot_vj = uci_vj_tc + coef)

any(dat$lci_tot_vj > dat$tot_vj) # FALSE
any(dat$uci_tot_vj < dat$tot_vj) # FALSE

# Conditional probability
## condp scalar
dat %<>%
  mutate(lci_condp_scale = lci_tot_vj/lci_tau,
         condp_scale = tot_vj/tau, 
         uci_condp_scale = uci_tot_vj/uci_tau)

any(dat$lci_condp_scale > dat$condp_scale) # TRUE
any(dat$uci_condp_scale < dat$condp_scale) # TRUE

## condp numerator
dat %<>%
  mutate(lci_condp_num = exp(lci_condp_scale),
         condp_num = exp(condp_scale), 
         uci_condp_num = exp(uci_condp_scale))

any(dat$lci_condp_num > dat$condp_num) # TRUE
any(dat$uci_condp_num < dat$condp_num) # FALSE

## condp denominator
tmp <- dat %>% group_by(trip_id, cset) %>% summarise(lci_condp_dom = sum(lci_condp_num)) %>% ungroup()
dat %<>% left_join(tmp) 

tmp <- dat %>% group_by(trip_id, cset) %>% summarise(condp_dom = sum(condp_num)) %>% ungroup()
dat %<>% left_join(tmp) 

tmp <- dat %>% group_by(trip_id, cset) %>% summarise(uci_condp_dom = sum(uci_condp_num)) %>% ungroup()
dat %<>% left_join(tmp) 

any(dat$lci_condp_dom > dat$condp_dom) # TRUE
any(dat$uci_condp_dom < dat$condp_dom) # FALSE

# condp
dat %<>%
  mutate(lci_condp = lci_condp_num/lci_condp_dom,
         condp = condp_num/condp_dom, 
         uci_condp = uci_condp_num/uci_condp_dom)

any(dat$lci_condp > dat$condp) # TRUE
any(dat$uci_condp < dat$condp) # TRUE

# condp iv (inclusive value of condp)
dat %<>%
  mutate(lci_iv = log(lci_condp_dom),
         iv = log(condp_dom), 
         uci_iv = log(uci_condp_dom))

any(dat$lci_iv > dat$iv) # TRUE
any(dat$uci_iv < dat$iv) # FALSE

# P1: probability of chooses nest
## p1 scalar
dat %<>%
  mutate(lci_p1_scale = lci_iv*lci_tau,
         p1_scale = iv*tau,
         uci_p1_scale = uci_iv*uci_tau)

any(dat$lci_p1_scale > dat$p1_scale) # FALSE
any(dat$uci_p1_scale < dat$p1_scale) # FALSE

## p1 numerator
dat %<>%
  mutate(lci_p1_num = exp(lci_p1_scale),
         p1_num = exp(p1_scale), 
         uci_p1_num = exp(uci_p1_scale))

any(dat$lci_p1_num > dat$p1_num) # FALSE
any(dat$uci_p1_num < dat$p1_num) # FALSE

# p1 denominator
tmp <- dat %>% group_by(trip_id) %>% distinct(lci_p1_num) %>% summarise(lci_p1_dom = sum(lci_p1_num))
dat %<>% left_join(tmp) 

tmp <- dat %>% group_by(trip_id) %>% distinct(p1_num) %>% summarise(p1_dom = sum(p1_num))
dat %<>% left_join(tmp) 

tmp <- dat %>% group_by(trip_id) %>% distinct(uci_p1_num) %>% summarise(uci_p1_dom = sum(uci_p1_num))
dat %<>% left_join(tmp) 

any(dat$lci_p1_dom > dat$p1_dom) # FALSE
any(dat$uci_p1_dom < dat$p1_dom) # FALSE

# p1
dat %<>%
  mutate(lci_p1 = lci_p1_num/lci_p1_dom,
         p1 = p1_num/p1_dom,
         uci_p1 = uci_p1_num/uci_p1_dom)

any(dat$lci_p1 > dat$p1) # TRUE
any(dat$uci_p < dat$p1) # FALSE

## P2: pr that agent will choose bottom level alternative
dat %<>%
  mutate(lci_p2 = lci_condp * lci_p1,
         p2 = condp * p1, 
         uci_p2 = uci_condp * uci_p1)

any(dat$lci_p2 > dat$p2) # TRUE
any(dat$uci_p2 < dat$p2) # TRUE

# Logsum base
dat %<>%
  mutate(lci_logsum_base = log(lci_p1_dom),
         logsum_base = log(p1_dom), 
         uci_logsum_base = log(uci_p1_dom))

any(dat$lci_logsum_base > dat$logsum_base) # FALSE
any(dat$uci_logsum_base < dat$logsum_base) # FALSE

# dat[which(dat$trip_id == 259), c(1, 45, 213:214, 223, 228:length(dat))]
```

# Counterfactual 
```{r}
# Change utility to zero for fishers
dat$lci_vj_sim <- dat$lci_tot_vj
dat$vj_sim <- dat$tot_vj
dat$uci_vj_sim <- dat$uci_tot_vj

identical(dat$lci_vj_sim, dat$lci_tot_vj) # TRUE
identical(dat$vj_sim, dat$tot_vj) # TRUE
identical(dat$uci_vj_sim, dat$uci_tot_vj) # TRUE

# dat %<>% mutate(lci_vj_sim = ifelse(zone_rm == 1, 0, lci_vj_sim)) 
# dat %<>% mutate(vj_sim = ifelse(zone_rm == 1, 0, vj_sim)) 
# dat %<>% mutate(uci_vj_sim = ifelse(zone_rm == 1, 0, uci_vj_sim)) 

dat %<>% filter(zone_rm == 0)

identical(dat$lci_vj_sim, dat$lci_tot_vj) # FALSE
identical(dat$vj_sim, dat$tot_vj) # FALSE
identical(dat$uc_vj_sim, dat$uci_tot_vj) # FALSE

# Conditional probability
## condp scalar
dat %<>%
  mutate(lci_condp_scale_sim = lci_vj_sim/lci_tau,
         condp_scale_sim = vj_sim/tau, 
         uci_condp_scale_sim = uci_vj_sim/uci_tau)

any(dat$lci_condp_scale_sim > dat$condp_scale_sim) # TRUE
any(dat$uci_condp_scale_sim < dat$condp_scale_sim) # TRUE

## condp numerator
dat %<>%
  mutate(lci_condp_num_sim = exp(lci_condp_scale_sim),
         condp_num_sim = exp(condp_scale_sim), 
         uci_condp_num_sim = exp(uci_condp_scale_sim))

any(dat$lci_condp_num_sim > dat$condp_num_sim) # TRUE
any(dat$uci_condp_num_sim < dat$condp_num_sim) # TRUE

## condp denominator
tmp <- dat %>% group_by(trip_id, cset) %>% summarise(lci_condp_dom_sim = sum(lci_condp_num_sim)) %>% ungroup()
dat %<>% left_join(tmp) 

tmp <- dat %>% group_by(trip_id, cset) %>% summarise(condp_dom_sim = sum(condp_num_sim)) %>% ungroup()
dat %<>% left_join(tmp) 

tmp <- dat %>% group_by(trip_id, cset) %>% summarise(uci_condp_dom_sim = sum(uci_condp_num_sim)) %>% ungroup()
dat %<>% left_join(tmp) 

any(dat$lci_condp_dom_sim > dat$condp_dom_sim) # TRUE
any(dat$uci_condp_dom_sim < dat$condp_dom_sim) # TRUE

# condp
dat %<>%
  mutate(lci_condp_sim = lci_condp_num_sim/lci_condp_dom_sim,
         condp_sim = condp_num_sim/condp_dom_sim, 
         uci_condp_sim = uci_condp_num_sim/uci_condp_dom_sim)

any(dat$lci_condp_sim > dat$condp_sim) # TRUE
any(dat$uci_condp_sim < dat$condp_sim) # TRUE

# condp iv (inclusive value of condp)
dat %<>%
  mutate(lci_iv_sim = log(lci_condp_dom_sim),
         iv_sim = log(condp_dom_sim), 
         uci_iv_sim = log(uci_condp_dom_sim))

any(dat$lci_iv_sim > dat$iv_sim) # TRUE
any(dat$uci_iv_sim < dat$iv_sim) # TRUE

# P1: probability of chooses nest
## p1 scalar
dat %<>%
  mutate(lci_p1_scale_sim = lci_iv_sim*lci_tau,
         p1_scale_sim = iv_sim*tau,
         uci_p1_scale_sim = uci_iv_sim*uci_tau)

any(dat$lci_p1_scale_sim > dat$p1_scale_sim) # FALSE
any(dat$uci_p1_scale_sim < dat$p1_scale_sim) # FALSE

## p1 numerator
dat %<>%
  mutate(lci_p1_num_sim = exp(lci_p1_scale_sim),
         p1_num_sim = exp(p1_scale_sim), 
         uci_p1_num_sim = exp(uci_p1_scale_sim))

any(dat$lci_p1_num_sim > dat$p1_num_sim) # FALSE
any(dat$uci_p1_num_sim < dat$p1_num_sim) # FALSE

# p1 denominator
tmp <- dat %>% group_by(trip_id) %>% distinct(lci_p1_num_sim) %>% summarise(lci_p1_dom_sim = sum(lci_p1_num_sim))
dat %<>% left_join(tmp) 

tmp <- dat %>% group_by(trip_id) %>% distinct(p1_num_sim) %>% summarise(p1_dom_sim = sum(p1_num_sim))
dat %<>% left_join(tmp) 

tmp <- dat %>% group_by(trip_id) %>% distinct(uci_p1_num_sim) %>% summarise(uci_p1_dom_sim = sum(uci_p1_num_sim))
dat %<>% left_join(tmp) 

any(dat$lci_p1_dom_sim > dat$p1_dom_sim) # FALSE
any(dat$uci_p1_dom_sim < dat$p1_dom_sim) # FALSE

# p1
dat %<>%
  mutate(lci_p1_sim = lci_p1_num_sim/lci_p1_dom_sim,
         p1_sim = p1_num_sim/p1_dom_sim,
         uci_p1_sim = uci_p1_num_sim/uci_p1_dom_sim)

any(dat$lci_p1_sim > dat$p1_sim) # TRUE
any(dat$uci_p_sim < dat$p1_sim) # FALSE

## P2: pr that agent will choose bottom level alternative
dat %<>%
  mutate(lci_p2_sim = lci_condp_sim * lci_p1_sim,
         p2_sim = condp_sim * p1_sim, 
         uci_p2_sim = uci_condp_sim * uci_p1_sim)

any(dat$lci_p2_sim > dat$p2_sim) # TRUE
any(dat$uci_p2_sim < dat$p2_sim) # TRUE

# Logsum base
dat %<>%
  mutate(lci_logsum_sim = log(lci_p1_dom_sim),
         logsum_sim = log(p1_dom_sim), 
         uci_logsum_sim = log(uci_p1_dom_sim))

any(dat$lci_logsum_sim > dat$logsum_sim) # FALSE
any(dat$uci_logsum_sim < dat$logsum_sim) # FALSE

dat[which(dat$trip_id == 259), c(1, 45, 213:214, 223, 229:231, 274:length(dat))]
```

# Welfare estimate
```{r}
b_fishers <- dat %>% filter(trip_id %in% fishers$trip_id)

tmp <- dat %>% group_by(trip_id) %>% summarise(logsum_base2 = sum(p1_dom))
tmp %<>% mutate(logsum_base2 = log(logsum_base2)) 
dat %<>% left_join(tmp) 

tmp <- dat %>% group_by(trip_id) %>% summarise(logsum_sim2 = sum(p1_dom_sim))
tmp %<>% mutate(logsum_sim2 = log(logsum_sim2)) 
dat %<>% left_join(tmp) 

a <- dat %>% group_by(trip_id) %>% 
  summarise(sub = logsum_sim2 - logsum_base2,
            w = sub/b$coef[1]) 
  
b$coef[1]

a %<>% distinct(trip_id, .keep_all = TRUE) 

a %<>% mutate(fisher = ifelse(trip_id %in% fishers$trip_id, 1, 0)) 
a %<>% group_by(fisher) %>% summarise(mean = mean(w))

# mean(a$w)
# 
# b_fishers %<>% 
#   mutate(av_lci_w = mean(lci_w),
#          av_w = mean(w),
#          av_uci_w = mean(uci_w))
#   
# any(dat$av_lci_w > dat$av_w) # FALSE
# any(dat$av_uci_w < dat$av_w) # FALSE
  
## Welfare estimates
# cs <- cbind(b_fishers[1, "av_lci_w"], b_fishers[1, "av_w"], b_fishers[1, "av_uci_w"])
# cs


# dat[which(dat$trip_id == 259), c(1, 45, 213:214, 223, 271:273, 307:length(dat))]
# 
# dat %>% filter(zone_rm == 1) %>% group_by(gridid_alt) %>% summarise(n=n())
# 
# which(dat$choice == 1 & dat$cset %in% "s_n" & dat$boat_access %in% "Yes")
# 
# dat[18800,]
# 
# dat[which(dat$trip_id == 279), c("cset", "gridid_alt","choice", "zone_rm","tot_vj", "vj_sim", "boat_access")]
```

# ARCHIVE
## Base without error
```{r}
# tmp <- dat %>% 
#   group_by(trip_id) %>% 
#   summarise(lci_logsum_base = mean(lci_logsum_base),
#             mean_logsum_base = mean(logsum_base),
#             uci_logsum_base = mean(uci_logsum_base))

# dat$vj_tc <- dat$fcflt_spgam * dat$tc_coef # utility of tc
# dat$vj_tot <- dat$vj_tc + dat$Coef # total utility

# Conditional probability
# dat$condp_scale <- dat$vj_tot/dat$tau 
# dat$condp_num <- exp(dat$condp_scale)
# tmp <- dat %>% group_by(trip_id, cset) %>% summarise(condp_dom = sum(condp_num)) %>% ungroup()
# dat %<>% left_join(tmp) 
# dat$condp <- dat$condp_num/dat$condp_dom
# dat$iv <- log(dat$condp_dom)

## P1: probability of choosen nest
# dat$p1_scale <- dat$iv*dat$tau
# Wnk: THIS IS WHERE YOU WOUDL ADD THE UTILITY FOR VARS THAT DESCRIBE YOUR NEST AND ADD IT TO p1_scale
# I do not have any in this model
# dat$p1_num <- exp(dat$p1_scale)
# tmp <- dat %>% group_by(trip_id) %>% distinct(p1_num) %>% summarise(p1_dom = sum(p1_num))
# dat %<>% left_join(tmp) 
# dat$p1 <- dat$p1_num/dat$p1_dom

# dat$p2 <- dat$condp * dat$p1

## logsum_base
# logsum_base <- log(dat$p1_dom)
```








